---
title: "Housekeeping P-P interactions"
author: "Martin Loza"
date: "2023/09/11
---

I would like to explore the existence of housekeeping promoter-promoter interactions. In other words, promoter-promoter interactions found in a large number of cell types.

```{r setup}
library(dplyr)
library(ggplot2)
library(patchwork)
library(stringr)
library(GenomicRanges)


## Global variables
in_dir  <- "/Volumes/MARTIN_LOZA/Projects/Enhancer_Grammar/Data/"
out_dir <- "~/Documents/Projects/Enhancer_Grammar/Analysis/Weekly_analyses/2023_09/Results"

date = 230911
red = "#CC6677"
blue = "#6699CC"
yellow = "#DDCC77"
dpi = 300
text_size = 14
fig_size = 4
point_size = 1
alpha = 0.3

#Functions
source(file = "~/Documents/Projects/Enhancer_Grammar/Analysis/Functions.R")
source(file = "~/Documents/Projects/Enhancer_Grammar/Analysis/Functions_Visualizations.r")

#Local functions 

```

## Load data

load cres with their predicted targets

```{r}
cres <- readRDS(file = "/Volumes/MARTIN_LOZA/Projects/Enhancer_Grammar/Data/v4/annotated/all_regions_annotated_230317.rds")
cres %>% filter(!duplicated(region_id)) %>% pull(n_celltypes) %>% table()
```
load the original regions containing the whole set of cell types

```{r}
regions_all <- readRDS(file = "/Volumes/MARTIN_LOZA/Projects/Enhancer_Grammar/Data/v4/raw_enhancers/merged/05_raw_enhancers_selected_50cts_221121.rds")
```

## Set up

Filter only the cres from selected cell types

```{r}
regions_all <- regions_all %>% filter(CellType %in% cres$celltypes)
```

```{r}
#select important columns 
cres <- cres %>% select(chr, start, end, region_id, center, length, celltypes, n_celltypes, test_n_celltypes, gene, tss, distance, n_genes, nearest_gene, nearest_transcript_id, nearest_transcript_tss, dist_to_nearest_transcript, ann_v1)
```

I would like to select interactions only for CP

```{r}
cres <- cres %>% filter(ann_v1 == "CP")
cres$ann_v1 <- droplevels(cres$ann_v1)
cres %>% filter(!duplicated(region_id)) %>% pull(n_celltypes) %>% table()
```
## Analysis

Let's rename CREs to interactions. 

```{r}
interactions_cp <- cres 
rm(cres)
gc()
#add new identifier with region_id and target gene
interactions_cp <- interactions_cp %>% mutate(id = paste0(region_id,"_", gene))
#test 
any(duplicated(interactions_cp$id))
```

To remove any bias added by a core-promoter predicted to interact with itself, let's remove CREs annotated that are predicted to interact with a target within 500 bps

```{r}
#remove cres_gene interactions of CP targetting itself withing 500 bps
# rmv_df <- interactions_cp %>% 
#   filter(gene == nearest_gene) %>% 
#   filter(distance <= 500)
# 
# #remove identified self-target from interactions df
# interactions_cp <- interactions_cp %>% filter(!id %in% rmv_df$id)
# 
# rm(rmv_df)
```

I would like count, for each interaction in how many cell types is found

For each interaction

For each cres, we would like to count the number of celltypes that a given target appears

```{r}
#get unique cres
u_cres <- interactions_cp %>% filter(!duplicated(region_id)) 
#for test
# u_cres <- u_cres %>% filter(test_n_celltypes == 50)
#for each unique cres
n_ct_interaction <- lapply(X = u_cres$region_id, FUN = function(c_region_id){
  #test
  # n_ct_interaction <- lapply(X = u_cres$region_id[1:10], FUN = function(c_region_id){
  c_region <- u_cres %>% filter(region_id == c_region_id)
  
  #get overlapping interactions in the original data set
  ovl_idx <- WhichRegionsOverlap(query_enhancer = c_region,
                                 subject_enhancer = regions_all)
    #get overlap regions from the original raw data
  all_overlaps <- regions_all[subjectHits(ovl_idx),]
  #count the number of times a target gene appear, e.g. how many cell types it appears as target
  c_n_cts <- all_overlaps %>% group_by(TargetGene) %>% count() %>%
    mutate(id = paste0(c_region$region_id,"_", TargetGene)) %>% ungroup() %>% select(id, n)
  
  return(c_n_cts)
})

n_ct_interaction <- Reduce(f = rbind, x = n_ct_interaction)

#save tmp results
saveRDS(object = n_ct_interaction, file = paste0(out_dir,"/tmp/n_ct_interactions_",date,".rds"))
#read tmp results
# n_ct_interaction <- readRDS(file = paste0(out_dir,"/tmp/n_ct_interactions_",date,".rds"))
```

## Test

### Test 1
Suspicious interactions

It looks like there are interactions that I didn't have before... I am not sure why...

```{r}
#short test
any(duplicated(n_ct_interaction$id))
any(duplicated(interactions_cp$id))

sum(interactions_cp$id %in% n_ct_interaction$id) == nrow(interactions_cp)
```

```{r}
suspicious_ids <- which(!n_ct_interaction$id %in% interactions_cp$id)
suspicious_ids <- n_ct_interaction[suspicious_ids,]
```

```{r}
interactions_cp %>% filter(id == "chr16:28823409-28823831_ATP2A1-AS1")
interactions_cp %>% filter(region_id == "chr16:28823409-28823831")
```

```{r}
suspicious_ids <- suspicious_ids %>%
  mutate(region_id = gsub(x = id, replacement = "", pattern = "_.*")) %>%mutate(gene = gsub(x = id, replacement = "", pattern = ".*_"))
```

### Test 2.

Let's test one of the regions with more than 50 counts

```{r}
n_ct_interaction <- n_ct_interaction %>% arrange(desc(n))
```

```{r}
test_interactions <- interactions_cp %>% filter(id == n_ct_interaction$id[1])
#get overlapping interactions in the original data set
ovl_idx <- WhichRegionsOverlap(query_enhancer = test_interactions,
                                 subject_enhancer = regions_all)
#get overlap regions from the original raw data
all_overlaps <- regions_all[subjectHits(ovl_idx),] %>% filter(TargetGene == test_interactions$gene) 
```

```{r}
all_overlaps %>% count(TargetGene, CellType) 
```

How to improve the count ...

```{r}
test_interactions <- interactions_cp %>% filter(id == n_ct_interaction$id[1])
#get overlapping interactions in the original data set
ovl_idx <- WhichRegionsOverlap(query_enhancer = test_interactions,
                                 subject_enhancer = regions_all)
#get overlap regions from the original raw data
all_overlaps <- regions_all[subjectHits(ovl_idx),]
all_overlaps %>% group_by(TargetGene) %>% count(CellType) %>% ungroup() %>% count(TargetGene)
```

```{r}
all_overlaps %>% filter(TargetGene == "GPKOW") %>% pull(CellType) %>% unique()
```

### Test 3.

Test unique cell types..

```{r}
sum(unique(regions_all$CellType) %in% interactions_cp$celltypes)
unique(regions_all$CellType)
```

