---
title: "Housekeeping P-P interactions"
author: "Martin Loza"
date: "2023/09/06
---

For the set of promoters and core-promoters, I would like to get mutual relatioships. For example, two promoters that mutually target each other.

NOTES:

- For now, let's analyse only the core promoter regions. Later, let's include the promoters too.
- For now, let's assume that the regions corresponds to an annotated gene
- Later, let's check for divergent genes too. For now, let's use the nearest genes

Notes:
I tried the next:

- For now, let's use a symmetric adjacency matrix. In this case the final count is kind of the total number of in/out edges.

```{r setup}
library(dplyr)
library(ggplot2)
library(patchwork)
library(stringr)
library(GenomicRanges)
library(igraph)
library(org.Hs.eg.db)
library(AnnotationDbi)
library(ggraph)

## Global variables
in_dir  <- "/Volumes/MARTIN_LOZA/Projects/Enhancer_Grammar/Data/"
out_dir <- "~/Documents/Projects/Enhancer_Grammar/Analysis/Weekly_analyses/2023_09/2023_09_12/Results/Mutual_P_P/"

date = 230906
red = "#CC6677"
blue = "#6699CC"
yellow = "#DDCC77"
dpi = 300
text_size = 14
fig_size = 4
point_size = 1
alpha = 0.3

#Functions
source(file = "~/Documents/Projects/Enhancer_Grammar/Analysis/Functions.R")
source(file = "~/Documents/Projects/Enhancer_Grammar/Analysis/Functions_Visualizations.r")

#Local functions 
GetGenesAnnotations <- function(keys = NULL, keys_type = "SYMBOL", annotations = c("SYMBOL","ENSEMBL","ENTREZID"), organism = "human"){
  if(organism == "human"){
    gene_annotations <- AnnotationDbi::select(x = org.Hs.eg.db, 
                                              keys = keys,
                                              columns = annotations,
                                              keytype = keys_type)
  }
  return(gene_annotations)
}
```

## Load data

load cres with their predicted targets

```{r}
cres <- readRDS(file = "/Volumes/MARTIN_LOZA/Projects/Enhancer_Grammar/Data/v4/annotated/all_regions_annotated_230317.rds")
cres %>% filter(!duplicated(region_id)) %>% pull(n_celltypes) %>% table()
```
## Set up

```{r}
#select important columns 
cres <- cres %>% dplyr::select(chr, start, end, region_id, center, length, celltypes, n_celltypes, test_n_celltypes, gene, tss, distance, nearest_gene, nearest_transcript_id, nearest_transcript_tss, dist_to_nearest_transcript, ann_v1)
```

I would like to select interactions only for CP or promoters
For now, let's do it only for core promoters

```{r}
# cres <- cres %>% filter(ann_v1 == "CP" | ann_v1 == "P")
cres <- cres %>% filter(ann_v1 == "CP")
cres$ann_v1 <- droplevels(cres$ann_v1)
cres %>% filter(!duplicated(region_id)) %>% pull(n_celltypes) %>% table()
```
## Analysis

Let's rename CREs to interactions. 

```{r}
interactions_cp <- cres 
rm(cres)
gc()
```

```{r}
#add new identifier with region_id and target gene
interactions_cp <- interactions_cp %>% mutate(id = paste0(region_id,"_", gene))
#test 
any(duplicated(interactions_cp$id))
```
To remove any bias added by a core-promoter predicted to interact with itself, let's remove CREs annotated that are predicted to interact with a target within 500 bps

```{r}
#remove cres_gene interactions of CP targetting itself withing 500 bps
rmv_df <- interactions_cp %>% 
  filter(gene == nearest_gene) %>% 
  filter(distance <= 500)

#remove identified self-target from interactions df
interactions_cp <- interactions_cp %>% filter(!id %in% rmv_df$id)

rm(rmv_df)
```

I would like get mutually target regions

For now, let's assume that the regions corresponds to an annotated gene

Later, let's check for divergent genes too. For now, let's use the nearest genes
```{r}
#set new genes with direction
interactions_cp <- interactions_cp %>% mutate(gene_from = nearest_gene, gene_to = gene)
#get the edges data frames
edges_df <- interactions_cp %>% dplyr::select(gene_from, gene_to)
```

### Symetric adjacency matrix

We count equally edges in/out. We can get the total counts per gene.

For this we need an indirected graph

```{r}
#get graph from edges
graph_interactions <- graph.data.frame(d = edges_df, directed = FALSE)
#get adjacency matrix
sym_matrix <- get.adjacency(graph_interactions)# %>% as.matrix()
sym_matrix[1:10, 1:10]
```

Let's get the degree of each gene
This will be the sum of rows
```{r}
degree <- igraph::degree(graph = graph_interactions, mode = "total", loops = TRUE, normalized = TRUE)
degree <- data.frame(gene = names(degree), degree = degree)
```

Let's check the genes with highest degree!!

```{r}
top_degree <- degree %>% slice_max(order_by = degree, n = 200)
top_degree
```

Check GO 

```{r}
symbol = GetGenesAnnotations(keys = top_degree$gene, keys_type = "SYMBOL")
write.csv(x = symbol[["ENTREZID"]],
          file = paste0(out_dir,"/degree_200_genes.csv"),
          row.names = FALSE, col.names = FALSE, quote = FALSE)
```

They are highly related to stress!! This align with previous knowledge.

### Genes with highly degree IN

We need a directed network for this.

```{r}
#get graph from edges
graph_directed <- graph.data.frame(d = edges_df, directed = TRUE)
```

```{r}
degree_in <- igraph::degree(graph = graph_directed, mode = "in", loops = TRUE, normalized = TRUE)
degree_in <- data.frame(gene = names(degree_in), degree = degree_in)
```

Let's check the genes with higher degree!!

```{r}
top_degree <- degree_in %>% slice_max(order_by = degree, n = 200)
top_degree
```

Check GO 

```{r}
symbol = GetGenesAnnotations(keys = top_degree$gene, keys_type = "SYMBOL")
write.csv(x = symbol[["ENTREZID"]],
          file = paste0(out_dir,"/degree_IN_200_genes.csv"),
          row.names = FALSE, col.names = FALSE, quote = FALSE)
```

### Genes with highly degree OUT

We need a directed network for this.

```{r}
#get graph from edges
graph_directed <- graph.data.frame(d = edges_df, directed = TRUE)
```

```{r}
degree_out <- igraph::degree(graph = graph_directed, mode = "out", loops = TRUE, normalized = TRUE)
degree_out <- data.frame(gene = names(degree_out), degree = degree_out)
```

Let's check the genes with higher degree!!

```{r}
top_degree <- degree_out %>% slice_max(order_by = degree, n = 200)
top_degree
```

Check GO 

```{r}
symbol = GetGenesAnnotations(keys = top_degree$gene, keys_type = "SYMBOL")
write.csv(x = symbol[["ENTREZID"]],
          file = paste0(out_dir,"/degree_OUT_200_genes.csv"),
          row.names = FALSE, col.names = FALSE, quote = FALSE)
```

### Plot

```{r}
#get graph from edges
graph_directed <- graph.data.frame(d = edges_df, directed = TRUE)
```

THIS IS TO COMPLEX. TOO MANY EDGES
```{r}
# ggraph(graph_directed, layout = 'kk') + 
#     geom_edge_fan(aes(alpha = after_stat(index)), show.legend = FALSE) + 
#     geom_node_point() + 
#     theme_graph(foreground = 'steelblue', fg_text_colour = 'white')
```

Let's try to select the top nodes

```{r}
degree_out <- igraph::degree(graph = graph_directed, mode = "all", loops = TRUE, normalized = TRUE)
degree_out <- data.frame(gene = names(degree_out), degree = degree_out)
top_degree <- degree_out %>% slice_max(order_by = degree, n = 3)
```

```{r}
test_df <- edges_df %>% filter(gene_from %in% top_degree$gene[1] | gene_to %in% top_degree$gene[1])
test_graph <- graph.data.frame(d = test_df, directed = TRUE)
```

```{r}
degree <- igraph::degree(graph = test_graph, mode = "all", loops = TRUE, normalized = TRUE)
d_sum <- summary(degree)
```

```{r}
thr <- d_sum["3rd Qu."]
ggraph(test_graph, layout = "centrality",cent=degree) + 
    geom_edge_link0( edge_alpha = 0.1) + 
    geom_node_point(aes(size = degree), color = "gray") + 
    theme_graph( fg_text_colour = 'white') + 
  geom_node_text(aes(label = name, filter = degree > thr), 
                 color = 'black', 
                 size = 3)
```

```{r}
test_df <- edges_df %>% filter(gene_from %in% top_degree$gene[2] | gene_to %in% top_degree$gene[2])
test_graph <- graph.data.frame(d = test_df, directed = TRUE)
```

```{r}
degree <- igraph::degree(graph = test_graph, mode = "all", loops = TRUE, normalized = TRUE)
d_sum <- summary(degree)
```

```{r}
thr <- d_sum["3rd Qu."]
ggraph(test_graph, layout = "centrality",cent=degree) + 
    geom_edge_link0(edge_alpha = 0.1) + 
    geom_node_point(aes(size = degree), color = "gray") + 
    theme_graph( fg_text_colour = 'white') + 
  geom_node_text(aes(label = name, filter = degree > thr), 
                 color = 'black', 
                 size = 3)
```

```{r}
test_df <- edges_df %>% filter(gene_from %in% top_degree$gene[3] | gene_to %in% top_degree$gene[3])
test_graph <- graph.data.frame(d = test_df, directed = TRUE)
```

```{r}
degree <- igraph::degree(graph = test_graph, mode = "all", loops = TRUE, normalized = TRUE)
d_sum <- summary(degree)
```

```{r}
thr <- d_sum["3rd Qu."]
ggraph(test_graph, layout = "centrality",cent=degree) + 
    geom_edge_link0( edge_alpha = 0.1) + 
    geom_node_point(aes(size = degree), color = "gray") + 
    theme_graph( fg_text_colour = 'white') + 
  geom_node_text(aes(label = name, filter = degree > thr), 
                 color = 'black', 
                 size = 3)
```

```{r}
test_df <- edges_df %>% filter(gene_from %in% top_degree$gene[1:3] | gene_to %in% top_degree$gene[1:3])
test_graph <- graph.data.frame(d = test_df, directed = TRUE)
```

```{r}
degree <- igraph::degree(graph = test_graph, mode = "all", loops = TRUE, normalized = TRUE)
d_sum <- summary(degree)
```

```{r}
thr <- d_sum["3rd Qu."]
ggraph(test_graph, layout = 'kk') + 
    geom_edge_link0( edge_alpha = 0.1) + 
    geom_node_point(aes(size = degree), color = "gray") + 
    theme_graph( fg_text_colour = 'white') + 
  geom_node_text(aes(label = name, filter = degree > thr), 
                 color = 'black', 
                 size = 3)
```

## Test

### Testing

Testing the in/out edges from the symmetric adjacency matrix of genes

```{r}
#this means tha GNB1 target itself 4 times. Let's verify this
tt <- interactions_cp %>% filter(gene_from == "GNB1" & gene_to == "CFAP74")
ttt <- interactions_cp %>% filter(gene_to == "GNB1" & gene_from == "CFAP74")
```



Let's try using directed graphs
```{r}
#get graph from edges
graph_interactions <- graph.data.frame(d = edges_df, directed = TRUE)
#get adjacency matrix
sym_matrix <- get.adjacency(graph_interactions)# %>% as.matrix()
sym_matrix[1:6, 1:6]
```

Let's count the number of interactions for each gene. We need to set as zero the upper triangle

```{r}
sym_matrix_zeros <- sym_matrix
sym_matrix_zeros[upper.tri(sym_matrix_zeros)] <- 0
sym_matrix_zeros[1:6, 1:6]
```
```{r}
total_interactions_row <- apply(X = sym_matrix_zeros, MARGIN = 1, FUN = sum)
total_interactions_row[1:6]
```

Let's count the number of interactions for each gene. We need to set as zero the upper triangle

```{r}
sym_matrix_zeros <- sym_matrix
sym_matrix_zeros[lower.tri(sym_matrix_zeros)] <- 0
sym_matrix_zeros[1:6, 1:6]
```

```{r}
total_interactions_col <- apply(X = sym_matrix_zeros, MARGIN = 2, FUN = sum)
total_interactions_col[1:6]
```

```{r}
total_interactions <- total_interactions_col + total_interactions_row
```

```{r}
total_interactions %>% sort(decreasing = TRUE)
```

```{r}
plot(x = graph_interactions)
```
```{r}
library(ggnetwork)
n <- ggnetwork(x = graph_interactions
```

```{r}
head(n)
```


```{r fig.width=10}
ggplot(n, aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_edges(color = "black") +
  # geom_nodelabel(aes(label = LETTERS[ name ]),
  #                fontface = "bold") +
  theme_blank()
```







## Tests

#### Test using indirected graph

It didn't work as I expected... 
The final count doesn't get what I want

```{r}
#get graph from edges
graph_interactions <- graph.data.frame(d = edges_df, directed = FALSE)
#get adjacency matrix
sym_matrix <- get.adjacency(graph_interactions) %>% as.matrix()
sym_matrix[1:6, 1:6]
```

Let's count the number of interactions for each gene. We need to set as zero the upper triangle

```{r}
sym_matrix_zeros <- sym_matrix
sym_matrix_zeros[upper.tri(sym_matrix_zeros)] <- 0
sym_matrix_zeros[1:6, 1:6]
```

```{r}
total_interactions <- apply(X = sym_matrix_zeros, MARGIN = 1, FUN = sum)
total_interactions[1:6]
```
```{r}
# Create a sample matrix
mat <- matrix(1:9, nrow = 3)

# Convert the upper triangle to zeros
mat[upper.tri(mat)] <- 0

# Print the modified matrix
print(mat)
```



## Tests

### Test for adjacency matrix

let's use the igraph package.

```{r}
# Create a simple undirected graph
edges <- data.frame(
  from = c("1", "2", "3", "4", "4"),
  to = c("2", "4", "1", "2", "2")
)

as.matrix(edges)

# Create a graph object
graph_directed <- graph.data.frame(edges, directed = TRUE)
graph_indirected <- graph.data.frame(edges, directed = FALSE)

# Get the adjacency matrices
sym_matrix <- get.adjacency(graph_indirected)
asym_matrix <- get.adjacency(graph_directed)

# sym_matrix
asym_matrix
sym_matrix
# as.matrix(sym_matrix) & as.matrix(asym_matrix)
```

I think I can use a symetric matrix from an indirected graph. Then count those interactions with more than one value 














adjacency matrix
```{r}
#get the universe of genes
genes_universe = union(interactions_cp$gene_enhancer, y = interactions_cp$gene_target)
#short test
any(duplicated(genes_universe))


```

```{r}
library(igraph)
```











For each interaction

```{r}
#init number of celltypes of interactions
n_ct_interaction <- matrix(data = 0, nrow = nrow(interactions_cp), ncol = 1)
#get cell types as a matrix to ease their index
n_celltypes_all <- matrix(regions_all$CellType, ncol = 1)

#for each interaction
# for(i in seq_len(nrow(interactions_cp))){
  for(i in 158426:nrow(interactions_cp)){
  #for tests
  # for(i in 1:10){
  # i = 1
  
  #get current interaction
  c_interaction = interactions_cp[i,]
  #get overlapping interactions in the original data set
  ovl_idx <- WhichRegionsOverlap(query_enhancer = c_interaction,
                                 subject_enhancer = regions_all)
  #get the current cell types
  c_ct <- n_celltypes_all[subjectHits(ovl_idx),1]
  #count the current cell types
  n_ct <- length(unique(c_ct))
  #assign the number of cell types
  n_ct_interaction[i,1] <- n_ct
  
}

#rmv unnecesary objects
rm( c_interaction, ovl_idx, c_ct, n_ct)
```




















How many CREs target each genes. I would like to know by cell types

```{r}
#count number of interactions
interactions_cp <- interactions_cp %>% 
  group_by(gene) %>%
  count() %>%
  right_join(.,y = interactions_cp, by = "gene")
#rename for downstream analyses
interactions_cp <- interactions_cp %>% 
  dplyr::rename(n_interactions = n)
```

Now I would like to know how many HK regions target each cp

```{r}
tmp_interactions <- interactions_cp %>% 
  group_by(gene) %>% 
  count(type == "HK")
#add colnames for further pipelines
colnames(tmp_interactions) <- c("gene", "is_HKG", "n_HK_CREs")
#select HK-interactions 
tmp_interactions <-tmp_interactions %>% filter(is_HKG == TRUE) %>% select(gene, n_HK_CREs)
#transfer information to main data frame
interactions_cp <- interactions_cp %>% left_join(y = tmp_interactions, by = "gene")
#Add zeros to NAs
idx <- which(is.na(interactions_cp$n_HK_CREs))
interactions_cp$n_HK_CREs[idx] <- 0

rm(tmp_interactions, idx)
```

Now, I would like to obtain a percentage of number of HK involved on each cp_gene

```{r}
interactions_cp <- interactions_cp %>% mutate(per_HK_CREs = n_HK_CREs/n_interactions)
```

Now I would like to know how many CTS regions target each cp

```{r}
tmp_interactions <- interactions_cp %>% 
  group_by(gene) %>% 
  count(n_celltypes == 1)
#add colnames for further pipelines
colnames(tmp_interactions) <- c("gene", "is_CTS", "n_CTS_CREs")
#select CTS-interactions 
tmp_interactions <-tmp_interactions %>% filter(is_CTS == TRUE) %>% select(gene, n_CTS_CREs)
#transfer information to main data frame
interactions_cp <- interactions_cp %>% left_join(y = tmp_interactions, by = "gene")
#Add zeros to NAs
idx <- which(is.na(interactions_cp$n_CTS_CREs))
interactions_cp$n_CTS_CREs[idx] <- 0

rm(tmp_interactions, idx)
```

Now, I would like to obtain a percentage of number of CTS involved on each cp_gene

```{r}
interactions_cp <- interactions_cp %>% mutate(per_CTS_CREs = n_CTS_CREs/n_interactions)
```

I would like to annotate which gene is a HKG for further comparisons 

```{r}
interactions_cp <- interactions_cp %>% mutate(is_HKG = ifelse(test = gene %in% hkg$gene_name, yes = TRUE, no = FALSE))
```

I would like to annote which gene is found to have a housekeeping core promoters

```{r}
interactions_cp <- interactions_cp %>%
  mutate(is_HK_CP = ifelse(test = gene %in% hk_cp_ann$gene_name, yes = TRUE, no = FALSE))
```

I would like to annotate how many CP or P and how many ESD, ELD, or ESLD target each gene

```{r}
interactions_cp <- interactions_cp %>% 
  group_by(gene) %>%
  summarise(n_P = sum(ann_v1 == "CP" | ann_v1 == "P"),
            n_E = sum(ann_v1 == "ESD" | ann_v1 == "ELD"| ann_v1 == "ESLD")) %>%
  right_join(y = interactions_cp, by = "gene")

#proportion of enhancers and promoters
interactions_cp <- interactions_cp %>% mutate(p_E = n_E/n_interactions)
interactions_cp <- interactions_cp %>% mutate(p_P = n_P/n_interactions)
```

All together

```{r}
#create new label
interactions_cp <- interactions_cp %>% mutate(type_target = "Other")
idx <- which(interactions_cp$is_HK_CP)
interactions_cp$type_target[idx] <- "nonHKG"
idx <- which(interactions_cp$is_HKG)
interactions_cp$type_target[idx] <- "HKG"

interactions_cp <- interactions_cp %>% 
  mutate(type_target = factor(type_target, 
                              levels = c("Other", "HKG", "nonHKG") ))
table(interactions_cp$type_target)
```

```{r}
saveRDS(object = interactions_cp, file = paste0(out_dir,"/interactions_cp_",date,".rds"))
```


### Plots

```{r}
plot_df <- interactions_cp %>% filter(!duplicated(gene))
plot_df$test_n_celltypes <- factor(plot_df$test_n_celltypes, levels = 1:50)
plot_df$n_celltypes <- factor(plot_df$n_celltypes, levels = c(1:44,50))
table(plot_df$type_target)
```
### Number of interactions

```{r}
my_comparison <- list(c("Other", "HKG"), 
                      c("Other", "nonHKG"),
                      c("HKG", "nonHKG"))

p <- plot_df %>% 
  MyComparisonPlot(wt = "n_interactions",
                   compare_wt = "type_target", 
                   palette =  c("gray", red, blue), 
                   adjust = 1,
                   comparisons = my_comparison,
                   label = "p.signif",
                   vjust = 0.5,
                   box_width = 0.15) + 
  ylab(expression("Number of interactions")) + 
  NoLegend() +
  theme(axis.title.x = element_blank(), plot.title = element_blank()) +
  theme(text = element_text(size = text_size)) +
  geom_jitter(size = 0.1, color = "black", alpha = 0.05)

p

# ggsave(filename = paste0("n_interactions_", date, ".png"), plot = p, device = "png",
#        path = out_dir, width = (fig_size), height = fig_size, dpi = dpi )
# ggsave(filename = paste0("n_interactions_", date, ".pdf"), plot = p, device = "pdf",
#        path = out_dir, width = (fig_size), height = fig_size, dpi = dpi )

```

```{r}
#statistics other
plot_df %>% filter(type_target == "Other") %>% pull(n_interactions) %>% summary()
#statistics other
plot_df %>% filter(type_target == "HKG") %>% pull(n_interactions) %>% summary()
#statistics other
plot_df %>% filter(type_target == "nonHKG") %>% pull(n_interactions) %>% summary()
```
Genes with ubiquitously active promoter are predicted as to have more interactions

### Distance of interactions

```{r fig.width=5, fig.height=5}
my_comparison <- list(c("Other", "HKG"), 
                      c("Other", "nonHKG"),
                      c("HKG", "nonHKG"))

p <- plot_df %>% 
  MyComparisonPlot(wt = "distance",
                   compare_wt = "type_target", 
                   palette =  c("gray", red, blue), 
                   adjust = 3,
                   comparisons = my_comparison,
                   label = "p.signif",
                   vjust = 0.5,
                   box_width = 0.15) + 
  scale_y_continuous(trans = "log10") +
  ylab(expression("Distance")) + 
  NoLegend() +
  theme(axis.title.x = element_blank(), plot.title = element_blank()) +
  theme(text = element_text(size = text_size)) +
  geom_jitter(size = 0.1, color = "black", alpha = 0.05)

p

# ggsave(filename = paste0("n_interactions_", date, ".png"), plot = p, device = "png",
#        path = out_dir, width = (fig_size), height = fig_size, dpi = dpi )
# ggsave(filename = paste0("n_interactions_", date, ".pdf"), plot = p, device = "pdf",
#        path = out_dir, width = (fig_size), height = fig_size, dpi = dpi )

```

```{r}
#statistics other
plot_df %>% filter(type_target == "Other") %>% pull(distance) %>% summary()
#statistics other
plot_df %>% filter(type_target == "HKG") %>% pull(distance) %>% summary()
#statistics other
plot_df %>% filter(type_target == "nonHKG") %>% pull(distance) %>% summary()
```
Distance between cts-CREs and HK-CREs

```{r}
plot_df %>% 
  filter(n_celltypes == 1 | n_celltypes == 50) %>% 
  ggplot(mapping = aes(x = distance, fill = n_celltypes)) +
  geom_histogram(position = "dodge") + 
  scale_x_continuous(trans = "pseudo_log")
```


### Number and proportion of promoters

I would like to analyze this from the perspective of number of enhancers an number of promoters

```{r}
my_comparison <- list(c("Other", "HKG"), 
                      c("Other", "nonHKG"),
                      c("HKG", "nonHKG"))

p1 <- plot_df %>% 
  MyComparisonPlot(wt = "p_P",
                   compare_wt = "type_target", 
                   palette =  c("gray", red, blue), 
                   adjust = 1,
                   comparisons = my_comparison,
                   label = "p.signif",
                   vjust = 0.3,
                   box_width = 0.15) + 
  ylab(expression("Proportion of promoters")) + 
  NoLegend() +
  theme(axis.title.x = element_blank(), plot.title = element_blank()) +
  theme(text = element_text(size = text_size)) +
  geom_jitter(size = 0.1, color = "black", alpha = 0.05)

p2 <- plot_df %>% 
  MyComparisonPlot(wt = "n_P",
                   compare_wt = "type_target", 
                   palette =  c("gray", red, blue), 
                   adjust = 1,
                   comparisons = my_comparison,
                   label = "p.signif",
                   vjust = 0.3,
                   box_width = 0.15) + 
  ylab(expression("Number of promoters")) + 
  NoLegend() +
  theme(axis.title.x = element_blank(), plot.title = element_blank()) +
  theme(text = element_text(size = text_size)) +
  geom_jitter(size = 0.1, color = "black", alpha = 0.05)

p1 + p2

# ggsave(filename = paste0("n_interactions_", date, ".png"), plot = p, device = "png",
#        path = out_dir, width = (fig_size), height = fig_size, dpi = dpi )
# ggsave(filename = paste0("n_interactions_", date, ".pdf"), plot = p, device = "pdf",
#        path = out_dir, width = (fig_size), height = fig_size, dpi = dpi )

```

```{r}
my_comparison <- list(c("HKG", "nonHKG"))

p1 <- plot_df %>% 
  filter(type_target != "Other") %>% 
  MyComparisonPlot(wt = "p_P",
                   compare_wt = "type_target", 
                   palette =  c("gray", red, blue), 
                   adjust = 1,
                   comparisons = my_comparison,
                   label = "p.signif",
                   vjust = 0.3,
                   box_width = 0.15) + 
  ylab(expression("Proportion of promoters")) + 
  NoLegend() +
  theme(axis.title.x = element_blank(), plot.title = element_blank()) +
  theme(text = element_text(size = text_size)) +
  geom_jitter(size = 0.1, color = "black", alpha = 0.05)

p2 <- plot_df %>% 
  filter(type_target != "Other") %>% 
  MyComparisonPlot(wt = "n_P",
                   compare_wt = "type_target", 
                   palette =  c("gray", red, blue), 
                   adjust = 1,
                   comparisons = my_comparison,
                   label = "p.signif",
                   vjust = 0.3,
                   box_width = 0.15) + 
  ylab(expression("Number of promoters")) + 
  NoLegend() +
  theme(axis.title.x = element_blank(), plot.title = element_blank()) +
  theme(text = element_text(size = text_size)) +
  geom_jitter(size = 0.1, color = "black", alpha = 0.05)

p1 + p2

```

```{r}
#statistics other
plot_df %>% filter(type_target == "Other") %>% pull(p_P) %>% summary()
#statistics other
plot_df %>% filter(type_target == "HKG") %>% pull(p_P) %>% summary()
#statistics other
plot_df %>% filter(type_target == "nonHKG") %>% pull(p_P) %>% summary()
```

```{r}
#statistics other
plot_df %>% filter(type_target == "HKG") %>% pull(n_P) %>% summary()
#statistics other
plot_df %>% filter(type_target == "nonHKG") %>% pull(n_P) %>% summary()
```

### Number and proportion of Enhancers

```{r}
my_comparison <- list(c("HKG", "nonHKG"))

p1 <- plot_df %>% 
  MyComparisonPlot(wt = "p_E",
                   compare_wt = "type_target", 
                   palette =  c("gray", red, blue), 
                   adjust = 1,
                   comparisons = my_comparison,
                   label = "p.signif",
                   vjust = 0.3,
                   box_width = 0.15) + 
  ylab(expression("Proportion of enhancers")) + 
  NoLegend() +
  theme(axis.title.x = element_blank(), plot.title = element_blank()) +
  theme(text = element_text(size = text_size)) +
  geom_jitter(size = 0.1, color = "black", alpha = 0.05)

p2 <- plot_df %>% 
  MyComparisonPlot(wt = "n_E",
                   compare_wt = "type_target", 
                   palette =  c("gray", red, blue), 
                   adjust = 1,
                   comparisons = my_comparison,
                   label = "p.signif",
                   vjust = 0.3,
                   box_width = 0.15) + 
  ylab(expression("Number of enhancers")) + 
  NoLegend() +
  theme(axis.title.x = element_blank(), plot.title = element_blank()) +
  theme(text = element_text(size = text_size)) +
  geom_jitter(size = 0.1, color = "black", alpha = 0.05)

p1 + p2

```

```{r}
my_comparison <- list(c("HKG", "nonHKG"))

p1 <- plot_df %>% 
  filter(type_target != "Other") %>% 
  MyComparisonPlot(wt = "p_E",
                   compare_wt = "type_target", 
                   palette =  c("gray", red, blue), 
                   adjust = 1,
                   comparisons = my_comparison,
                   label = "p.signif",
                   vjust = 0.3,
                   box_width = 0.15) + 
  ylab(expression("Proportion of enhancers")) + 
  NoLegend() +
  theme(axis.title.x = element_blank(), plot.title = element_blank()) +
  theme(text = element_text(size = text_size)) +
  geom_jitter(size = 0.1, color = "black", alpha = 0.05)

p2 <- plot_df %>% 
  filter(type_target != "Other") %>% 
  MyComparisonPlot(wt = "n_E",
                   compare_wt = "type_target", 
                   palette =  c("gray", red, blue), 
                   adjust = 1,
                   comparisons = my_comparison,
                   label = "p.signif",
                   vjust = 0.3,
                   box_width = 0.15) + 
  ylab(expression("Number of enhancers")) + 
  NoLegend() +
  theme(axis.title.x = element_blank(), plot.title = element_blank()) +
  theme(text = element_text(size = text_size)) +
  geom_jitter(size = 0.1, color = "black", alpha = 0.05)

p1 + p2

```

```{r}
#statistics other
plot_df %>% filter(type_target == "Other") %>% pull(p_E) %>% summary()
#statistics other
plot_df %>% filter(type_target == "HKG") %>% pull(p_E) %>% summary()
#statistics other
plot_df %>% filter(type_target == "nonHKG") %>% pull(p_E) %>% summary()
```

```{r}
#statistics other
plot_df %>% filter(type_target == "HKG") %>% pull(n_E) %>% summary()
#statistics other
plot_df %>% filter(type_target == "nonHKG") %>% pull(n_E) %>% summary()
```

### Number of enhancers and number of promoters 

```{r}
#housekeeping core promoters
plot_df %>% filter(is_HK_CP) %>%
  ggplot(mapping = aes(y = n_P, x = n_E)) +
  geom_point(alpha = 0.1)+ coord_fixed(xlim = c(0,60), ylim = c(0,60)) + theme_light() + ggtitle("HK-CPs")
#non-housekeeping core promoters
plot_df %>% filter(!is_HK_CP) %>%
  ggplot(mapping = aes(y = n_P, x = n_E)) +
  geom_point(alpha = 0.1) + coord_fixed(xlim = c(0,60), ylim = c(0,60))
#housekeeping genes
plot_df %>% filter(is_HKG) %>%
  ggplot(mapping = aes(y = n_P, x = n_E)) +
  geom_point(alpha = 0.1) + coord_fixed(xlim = c(0,60), ylim = c(0,60))
#housekeeping core promoters, except housekeeping genes
plot_df %>% filter(is_HK_CP) %>%
  filter(!is_HKG) %>%
  ggplot(mapping = aes(y = n_P, x = n_E)) +
  geom_point(alpha = 0.1) + coord_fixed(xlim = c(0,60), ylim = c(0,60))
```

```{r}
tmp_df <- plot_df %>% mutate(combination = paste(n_P, n_E, sep = "_"))
frequency_others <- tmp_df %>% group_by(combination) %>% count() %>% arrange(desc(n)) %>% rename(counts = n )
most_frequent <- frequency_others %>% ungroup() %>% arrange(desc(counts)) %>% slice_max(counts)

tmp_df %>% 
  left_join(., y = frequency_others, by = "combination") %>%
  ggplot(mapping = aes(x = n_P, y = n_E, color = counts)) +
  geom_point() +
  coord_fixed(xlim = c(0,60), ylim = c(0,60)) + theme_light() + ggtitle("All genes", subtitle = most_frequent$combination) + scale_color_gradient2( low= blue, mid = "gray", high=red)
```

```{r}
tmp_df <- plot_df %>% filter(is_HK_CP) %>% mutate(combination = paste(n_P, n_E, sep = "_"))
frequency_others <- tmp_df %>% group_by(combination) %>% count() %>% arrange(desc(n)) %>% rename(counts = n )
most_frequent <- frequency_others %>% ungroup() %>% arrange(desc(counts)) %>% slice_max(counts)

tmp_df %>% 
  left_join(., y = frequency_others, by = "combination") %>%
  ggplot(mapping = aes(x = n_P, y = n_E, color = counts)) +
  geom_point() +
  coord_fixed(xlim = c(0,60), ylim = c(0,60)) + theme_light() + ggtitle("HK-CPs", subtitle = most_frequent$combination) + scale_color_gradient2( low= blue, mid = "gray", high=red)
```

```{r}
#non-housekeeping core promoters
tmp_df <- plot_df %>% filter(!is_HK_CP) %>% mutate(combination = paste(n_P, n_E, sep = "_"))
frequency_others <- tmp_df %>% group_by(combination) %>% count() %>% arrange(desc(n)) %>% rename(counts = n )
most_frequent <- frequency_others %>% ungroup() %>% arrange(desc(counts)) %>% slice_max(counts)

tmp_df %>% 
  left_join(., y = frequency_others, by = "combination") %>%
  ggplot(mapping = aes(x = n_P, y = n_E, color = counts)) +
  geom_point() +
  coord_fixed(xlim = c(0,60), ylim = c(0,60)) + theme_light() + ggtitle("nonHK-CPs", subtitle = most_frequent$combination) + scale_color_gradient2( low= blue, mid = "gray", high=red)
```

```{r}
tmp_df <- plot_df %>% filter(is_HKG) %>% mutate(combination = paste(n_P, n_E, sep = "_"))
frequency_others <- tmp_df %>% group_by(combination) %>% count() %>% arrange(desc(n)) %>% rename(counts = n )
most_frequent <- frequency_others %>% ungroup() %>% arrange(desc(counts)) %>% slice_max(counts)

tmp_df %>% 
  left_join(., y = frequency_others, by = "combination") %>%
  ggplot(mapping = aes(x = n_P, y = n_E, color = counts)) +
  geom_point() +
  coord_fixed(xlim = c(0,60), ylim = c(0,60)) + theme_light() + ggtitle("HKG", subtitle = most_frequent$combination) + scale_color_gradient2( low= blue, mid = "gray", high=red)
```

```{r}
tmp_df <- plot_df %>% filter(is_HK_CP) %>%
  filter(!is_HKG) %>% mutate(combination = paste(n_P, n_E, sep = "_"))
frequency_others <- tmp_df %>% group_by(combination) %>% count() %>% arrange(desc(n)) %>% rename(counts = n )
most_frequent <- frequency_others %>% ungroup() %>% arrange(desc(counts)) %>% slice_max(counts)

tmp_df %>% 
  left_join(., y = frequency_others, by = "combination") %>%
  ggplot(mapping = aes(x = n_P, y = n_E, color = counts)) +
  geom_point() +
  coord_fixed(xlim = c(0,60), ylim = c(0,60)) + theme_light() + ggtitle("nonHKG", subtitle = most_frequent$combination) + scale_color_gradient2( low= blue, mid = "gray", high=red)
```


```{r}
#housekeeping core promoters
plot_df %>% filter(is_HK_CP) %>%
  ggplot(mapping = aes(y = p_P, x = p_E)) +
  geom_point(alpha = 0.1) + coord_fixed(xlim = c(0,1), ylim = c(0,1))
#non-housekeeping core promoters
plot_df %>% filter(!is_HK_CP) %>%
  ggplot(mapping = aes(y = p_P, x = p_E)) +
  geom_point(alpha = 0.1) + coord_fixed(xlim = c(0,1), ylim = c(0,1))
#housekeeping genes
plot_df %>% filter(is_HKG) %>%
  ggplot(mapping = aes(y = p_P, x = p_E)) +
  geom_point(alpha = 0.1) + coord_fixed(xlim = c(0,1), ylim = c(0,1))
#housekeeping core promoters, except housekeeping genes
plot_df %>% filter(is_HK_CP) %>%
  filter(!is_HKG) %>%
  ggplot(mapping = aes(y = p_P, x = p_E)) +
  geom_point(alpha = 0.1) + coord_fixed(xlim = c(0,1), ylim = c(0,1))
```

### Number of HK-CPs

```{r}
my_comparison <- list(c("Other", "HKG"), 
                      c("Other", "nonHKG"),
                      c("HKG", "nonHKG"))

p1 <- plot_df %>% 
  # filter(type_target != "Other") %>% 
  MyComparisonPlot(wt = "per_HK_CREs", 
                   compare_wt = "type_target", 
                   palette =  c("gray", red, blue), 
                   adjust = 1,
                   comparisons = my_comparison,
                   label = "p.signif",
                   vjust = 0.3,
                   box_width = 0.15) + 
  ylab(expression("Proportion of HK-CREs")) + 
  NoLegend() +
  theme(axis.title.x = element_blank(), plot.title = element_blank()) +
  theme(text = element_text(size = text_size)) +
  geom_jitter(size = 0.1, color = "black", alpha = 0.05)

p2 <- plot_df %>% 
  # filter(type_target != "Other") %>% 
  MyComparisonPlot(wt = "n_HK_CREs", 
                   compare_wt = "type_target", 
                   palette =  c("gray", red, blue), 
                   adjust = 1,
                   comparisons = my_comparison,
                   label = "p.signif",
                   vjust = 0.3,
                   box_width = 0.15) + 
  ylab(expression("Number of HK-CREs")) + 
  NoLegend() +
  theme(axis.title.x = element_blank(), plot.title = element_blank()) +
  theme(text = element_text(size = text_size)) +
  geom_jitter(size = 0.1, color = "black", alpha = 0.05)

p1 + p2

```

```{r}
#statistics other
plot_df %>% filter(type_target == "Other") %>% pull(n_HK_CREs) %>% summary()
#statistics HKG
plot_df %>% filter(type_target == "HKG") %>% pull(n_HK_CREs) %>% summary()
#statistics other
plot_df %>% filter(type_target == "nonHKG") %>% pull(n_HK_CREs) %>% summary()
```

```{r}
#statistics other
plot_df %>% filter(type_target == "Other") %>% pull(per_HK_CREs) %>% summary()
#statistics HKG
plot_df %>% filter(type_target == "HKG") %>% pull(per_HK_CREs) %>% summary()
#statistics other
plot_df %>% filter(type_target == "nonHKG") %>% pull(per_HK_CREs) %>% summary()
```

### Number of CTS-CPs

```{r}
my_comparison <- list(c("Other", "HKG"), 
                      c("Other", "nonHKG"),
                      c("HKG", "nonHKG"))

p1 <- plot_df %>% 
  MyComparisonPlot(wt = "per_CTS_CREs", 
                   compare_wt = "is_HK_CP", 
                   palette =  c("gray", red, blue), 
                   adjust = 2.5,
                   comparisons = my_comparison,
                   label = "p.signif",
                   vjust = 0.3,
                   box_width = 0.15) + 
  ylab(expression("Proportion of CTS-CREs")) + 
  NoLegend() +
  theme(axis.title.x = element_blank(), plot.title = element_blank()) +
  theme(text = element_text(size = text_size)) +
  geom_jitter(size = 0.1, color = "black", alpha = 0.05)

p2 <- plot_df %>% 
  MyComparisonPlot(wt = "n_CTS_CREs",  
                   compare_wt = "is_HK_CP", 
                   palette =  c("gray", red, blue), 
                   adjust = 2.5,
                   comparisons = my_comparison,
                   label = "p.signif",
                   vjust = 0.3,
                   box_width = 0.15) + 
  ylab(expression("Number of CTS-CREs")) + 
  NoLegend() +
  theme(axis.title.x = element_blank(), plot.title = element_blank()) +
  theme(text = element_text(size = text_size)) +
  geom_jitter(size = 0.1, color = "black", alpha = 0.05)

p1 + p2
```


```{r}
#statistics other
plot_df %>% filter(type_target == "Other") %>% pull(n_CTS_CREs) %>% summary()
#statistics HKG
plot_df %>% filter(type_target == "HKG") %>% pull(n_CTS_CREs) %>% summary()
#statistics other
plot_df %>% filter(type_target == "nonHKG") %>% pull(n_CTS_CREs) %>% summary()
```

```{r}
#statistics other
plot_df %>% filter(type_target == "Other") %>% pull(per_CTS_CREs) %>% summary()
#statistics HKG
plot_df %>% filter(type_target == "HKG") %>% pull(per_CTS_CREs) %>% summary()
#statistics other
plot_df %>% filter(type_target == "nonHKG") %>% pull(per_CTS_CREs) %>% summary()
```










