---
title: "P-P Interactions, analysis by graph, LOCAL"
author: "Martin Loza"
date: "2023/09/18
---

For the set of core-promoters, I would like to make a graph analysis of thein interactions. I would like to use the n_ct_interactions as a weight for the edges.

In previous analysis I obtained nodes for a long distance, I would like to set some constraints to avoid this poblem.


NOTES:


```{r setup}
library(dplyr)
library(ggplot2)
library(patchwork)
library(stringr)
library(GenomicRanges)
library(igraph)
library(ggraph)

## Global variables
in_dir  <- "~/Documents/Projects/HK_Interactions/Analysis/2023_09/2023_09_18/Results/"
out_dir <- "~/Documents/Projects/HK_Interactions/Analysis/2023_09/2023_09_18/Results/"

date = 2309018
red = "#CC6677"
blue = "#6699CC"
yellow = "#DDCC77"

#Functions
source(file = "~/Documents/Projects/HK_Interactions/Analysis/Functions.R")
source(file = "~/Documents/Projects/HK_Interactions/Analysis/Functions_Visualizations.R")

#Local functions 

GetGenerationEdges <- function(edges = NULL, init_nodes = NULL, n_generations = 2){
  
  #loop. In each generation get the edges associated to init nodes
  for(i in 2:n_generations){
    tmp_df <- edges %>% filter(gene_from %in% all_of(init_nodes) | gene_to %in% all_of(init_nodes))
    init_nodes <- union(tmp_df$gene_from, tmp_df$gene_to)
  }

  return(tmp_df)
}
```

## Load data

Load annotated interactions 

```{r}
interactions <- readRDS(paste0(in_dir,"/interactions_all_230912.Rds"))
interactions %>% filter(!duplicated(region_id)) %>% pull(n_celltypes) %>% table()
```

## Set up

I would like to select interactions only for CP 

```{r}
interactions <- interactions %>% filter(ann_v1 == "CP")
interactions$ann_v1 <- droplevels(interactions$ann_v1)
interactions %>% filter(!duplicated(region_id)) %>% pull(n_celltypes) %>% table()
```

Check for duplicated interactions 

```{r}
any(duplicated(interactions$id))
```

Set the absolute distance of interactions 

```{r}
interactions <- interactions %>%
  mutate(abs_distance = abs(distance))
```

## Analysis

### Set HK-CP

```{r}
#get HK-CP
hk_cp <- interactions %>% filter(n_celltypes == 50) %>% filter(!duplicated(region_id)) %>% select(chr, center)
#set fixed length regions
l = 500
hk_cp <- hk_cp %>% mutate(start = center - l, end = center + l)
```

create and save the bedpe

```{r}
tmp_bedpe <- hk_cp %>% mutate(chr1 = chr, chr2 = chr, x1 = start, x2 = end, y1 = start, y2 = end,
                        name = ".", score = ".", 
                        strand1 = ".", strand2 = ".") %>% 
  select(chr1, x1, x2, chr2, y1, y2, name, score, strand1, strand2)

#save bedpe
write.table(x = tmp_bedpe,
            file = "~/Documents/Projects/HK_Interactions/Analysis/2023_10/2023_10_02/tmp_data/hk_cp_raw.bedpe",
            quote = FALSE, sep = "\t", row.names = FALSE, col.names = FALSE)
```

Get HK_PPI

```{r}
max_dist = 2e6
hk_ppi <- interactions %>% filter(abs_distance <= max_dist) %>% filter(n_ct_interactions >= 45)
hk_ppi <- hk_ppi %>% mutate(interaction_id = paste0(nearest_gene,":", gene))
```

Prepare bed 

```{r}
bed_domains <- hk_ppi %>% select(chr, center, tss, interaction_id)
```

We need to be sure that the start is lower than the end

```{r}
bed_domains <- bed_domains %>% mutate(start = ifelse(center <= tss, center, tss)) %>% 
  mutate(end = ifelse(center > tss, center, tss)) %>% 
  select(chr, start, end, interaction_id)
```

```{r}
#short test
sum(bed_domains$start <= bed_domains$end) == nrow(bed_domains)
```

Filter regions with same start and end

```{r}
bed_domains <- bed_domains %>% filter(start != end)
```

```{r}
#short test
sum(bed_domains$start < bed_domains$end) == nrow(bed_domains)
```

save bed

```{r}
write.table(x = bed_domains, file = "~/Documents/Projects/HK_Interactions/Analysis/2023_10/2023_10_02/tmp_data/hk_ppi_domains.bed", quote = FALSE, sep = "\t", row.names = FALSE, col.names = FALSE)
```

## bedpe for Juicebox

```{r}
bedpe <- bed_domains %>% mutate(chr1 = chr, chr2 = chr, x1 = start, x2 = end, y1 = start, y2 = end,
                        name = ".", score = ".", 
                        strand1 = ".", strand2 = ".") %>% select(chr1, x1, x2, chr2, y1, y2, name, score, strand1, strand2)
```

save bedpe

```{r}
write.table(x = bedpe, file = "~/Documents/Projects/HK_Interactions/Analysis/2023_10/2023_10_02/tmp_data/hk_ppi_domains.bedpe", quote = FALSE, sep = "\t", row.names = FALSE, col.names = FALSE)
```

## Save location of HK-CP

```{r}
table(interactions)
```


```{r}
hk_cp_bedpe <- hk_ppi %>% select(chr, start, end) %>% mutate(chr1 = chr, chr2 = chr, x1 = start, x2 = end, y1 = start, y2 = end,
                        name = ".", score = ".", 
                        strand1 = ".", strand2 = ".") %>% select(chr1, x1, x2, chr2, y1, y2, name, score, strand1, strand2)
```

```{r}
write.table(x = hk_cp_bedpe, file = "~/Documents/Projects/HK_Interactions/Analysis/2023_10/2023_10_02/tmp_data/hk_cp.bedpe", quote = FALSE, sep = "\t", row.names = FALSE, col.names = FALSE)
```










