---
title: "Housekeeping P-P interactions"
author: "Martin Loza"
date: "2023/09/12
---

I would like to explore the existence of housekeeping promoter-promoter interactions. In other words, promoter-promoter interactions found in a large number of cell types.

In a previous version I found a bug that could produce counts higher than 50 (cell types) because of multiple overlapping regions from the cell type. I would like to correct this bug and re-run the analysis.

```{r setup}
library(dplyr)
library(ggplot2)
library(patchwork)
library(stringr)
library(GenomicRanges)


## Global variables
in_dir  <- "/Volumes/MARTIN_LOZA/Projects/Enhancer_Grammar/Data/"
out_dir <- "~/Documents/Projects/Enhancer_Grammar/Analysis/Weekly_analyses/2023_09/2023_09_18//Results"

date = 230912
red = "#CC6677"
blue = "#6699CC"
yellow = "#DDCC77"
dpi = 300
text_size = 14
fig_size = 4
point_size = 1
alpha = 0.3

#Functions
source(file = "~/Documents/Projects/Enhancer_Grammar/Analysis/Functions.R")
source(file = "~/Documents/Projects/Enhancer_Grammar/Analysis/Functions_Visualizations.r")

#Local functions 

```

## Load data

load cres with their predicted targets

```{r}
cres <- readRDS(file = "/Volumes/MARTIN_LOZA/Projects/Enhancer_Grammar/Data/v4/annotated/all_regions_annotated_230317.rds")
cres %>% filter(!duplicated(region_id)) %>% pull(n_celltypes) %>% table()
```
load the original regions containing the whole set of cell types

```{r}
regions_all <- readRDS(file = "/Volumes/MARTIN_LOZA/Projects/Enhancer_Grammar/Data/v4/raw_enhancers/merged/05_raw_enhancers_selected_50cts_221121.rds")
```

load genes universe to add gene information

```{r}
genes <- readRDS("/Volumes/MARTIN_LOZA/Projects/Enhancer_Grammar/Data/Commonly_used/hg38/hg38_genes_universe_221121.rds")
```

## Set up

Filter only the cres from selected cell types

```{r}
regions_all <- regions_all %>% filter(CellType %in% cres$celltypes)
```

```{r}
#select important columns 
cres <- cres %>% dplyr::select(chr, start, end, region_id, center, length, celltypes, n_celltypes, test_n_celltypes, gene, tss, distance, n_genes, nearest_gene, nearest_transcript_id, nearest_transcript_tss, dist_to_nearest_transcript, ann_v1)
```

I would like to select interactions only for CP

```{r}
# cres <- cres %>% filter(ann_v1 == "CP")
# cres$ann_v1 <- droplevels(cres$ann_v1)
# cres %>% filter(!duplicated(region_id)) %>% pull(n_celltypes) %>% table()
```
## Analysis

Let's rename CREs to interactions. 

```{r}
interactions_cp <- cres 
rm(cres)
gc()
#add new identifier with region_id and target gene
interactions_cp <- interactions_cp %>% mutate(id = paste0(region_id,"_", gene))
#test 
any(duplicated(interactions_cp$id))
```

I would like count, for each interaction in how many cell types is found

For each interaction

For each cres, we would like to count the number of celltypes that a given target appears

```{r}
#get unique cres
u_cres <- interactions_cp %>% filter(!duplicated(region_id)) 
#for test
# u_cres <- u_cres %>% filter(test_n_celltypes == 50)
#for each unique cres
# n_ct_interaction <- lapply(X = u_cres$region_id, FUN = function(c_region_id){
  # test
  n_ct_interaction <- lapply(X = u_cres$region_id[1:10], FUN = function(c_region_id){
  
  #test
  # c_region_id = u_cres$region_id[300]
  # c_region_id = "chr16:28823409-28823831"
  
  c_region <- u_cres %>% filter(region_id == c_region_id)
  
  #get overlapping interactions in the original data set
  ovl_idx <- WhichRegionsOverlap(query_enhancer = c_region,
                                 subject_enhancer = regions_all)
    #get overlap regions from the original raw data
  all_overlaps <- regions_all[subjectHits(ovl_idx),]
  #count the number of times a target gene appear, e.g. how many cell types it appears as target
  c_n_cts <- all_overlaps %>% group_by(TargetGene) %>%
    count(CellType) %>%
    ungroup() %>%
    count(TargetGene) %>%
    mutate(id = paste0(c_region$region_id,"_", TargetGene)) %>%
    select(id, n)  
  
  return(c_n_cts)
})

n_ct_interaction <- Reduce(f = rbind, x = n_ct_interaction)

#save tmp results
saveRDS(object = n_ct_interaction, file = paste0(out_dir,"/tmp/n_ct_interactions_v2_",date,".rds"))
#read tmp results
# n_ct_interaction <- readRDS(file = paste0(out_dir,"/tmp/n_ct_interactions_230911.rds"))
n_ct_interaction <- readRDS(file = paste0(out_dir,"/tmp/n_ct_all_interactions_v2_230912.rds"))
```

Prepare the data set. There are some extra interactions captured by this analysis. Let's try to include them.

```{r}
n_ct_interaction <- n_ct_interaction %>%
  mutate(gene = gsub(x = id, replacement = "", pattern =".*_")) %>%
  mutate(region_id = gsub(x = id, replacement = "", pattern ="_.*")) %>% dplyr::select(region_id, gene, n, id)
```

Let's merge the other information

Let's transfer the regions information

```{r}
tmp_df <- interactions_cp %>% filter(!duplicated(region_id)) %>%
  dplyr::select(region_id, chr, start, end,
         ann_v1, n_celltypes, test_n_celltypes, celltypes, 
         nearest_gene,
         nearest_transcript_tss, dist_to_nearest_transcript)
interactions_new <- n_ct_interaction %>% left_join(y = tmp_df, by = "region_id") 

#rename the column n and reorder the columns
interactions_new <- interactions_new %>% mutate("n_ct_interactions" = n) %>%
  dplyr::select(chr, start, end, region_id, ann_v1,
         gene, n_ct_interactions, id,
         n_celltypes, test_n_celltypes, celltypes, 
         nearest_gene,
         nearest_transcript_tss, dist_to_nearest_transcript )

#test
any(is.na(interactions_new))
```

Let's add the TSS information of genes

```{r}
#select info to add
tmp_genes <- genes %>% 
  select(gene_name, tss, strand) %>%
  mutate(gene_strand = strand, gene = gene_name) %>% 
  select(gene, tss, gene_strand)

#add info to interactions df
interactions_new <- interactions_new %>% left_join(y = tmp_genes, by = "gene") %>% #add target gene information
  mutate(length = end - start) %>% #add length and center of the regions
  mutate(center = start + floor(length/2)) %>%
  mutate(distance = tss - center) %>% #calculate the distance to the target genes
  select(chr, start, end, region_id, length, center, #select columns of interests
         ann_v1, n_celltypes, test_n_celltypes, 
         gene, tss, distance, gene_strand, 
         n_ct_interactions, celltypes,
         nearest_gene, nearest_transcript_tss, dist_to_nearest_transcript, id)
                                  
rm(tmp_genes)
```

## Quality control

```{r}
interactions_cp %>% 
  filter(!duplicated(region_id)) %>%
  pull(n_celltypes) %>% table()

interactions_new %>%
  filter(!duplicated(region_id)) %>% 
  pull(n_celltypes) %>% table()

```

### Remove suspicious regions

Let's remove suspicious regions which interact in a higher number of cell types than they are active in (See tests). This weird behavior was caused by the merging procedure. For now, we can filter them out.

```{r}
filter_regions <- interactions_new %>%
  filter(n_ct_interactions > n_celltypes) %>%
  filter(!duplicated(region_id))
#filter out suspicious regions 
interactions_new <- interactions_new %>%
  filter(!region_id %in% filter_regions$region_id)

#lets compare how many regions we lost
tmp_regions <- interactions_cp %>% filter(!duplicated(region_id)) %>% pull(n_celltypes)
tmp_2 <- interactions_new %>% filter(!duplicated(region_id)) %>% pull(n_celltypes)

table(tmp_regions) - table(tmp_2) 

cat("Number of regions: ", length(tmp_regions))

rm(filter_regions)
```
We didn't lose any cell type-specific nor hk regions

```{r}
interactions_new %>% filter(!duplicated(region_id)) %>% 
  ggplot(mapping = aes(x = n_celltypes)) + geom_histogram(stat = "count")
```

It looks good!. 

```{r}
interactions_cp %>% 
  filter(!duplicated(region_id)) %>%
  pull(n_celltypes) %>% table()

interactions_new %>%
  filter(!duplicated(region_id)) %>% 
  pull(n_celltypes) %>% table()

```

### Other QC

I THINK I CAN LEAVE THESE FOR NOW AS I MIGHT LOOSE SOME HK REGIONS THAT ONLY TARGET THEMSELVES. I WILL JUST KEEP IN MIND THAT INTERACTIONS WILL HAVE AN EXTRA +1 IN THE NUMBER 

To remove any bias added by a core-promoter predicted to interact with itself, let's remove CREs annotated that are predicted to interact with a target within 500 bps

```{r}
#remove cres_gene interactions of CP targetting itself withing 500 bps
# rmv_df <- interactions_new %>%
#   filter(gene == nearest_gene) %>%
#   filter(distance <= 500)
# 
# #remove identified self-target from interactions df
# interactions_new <- interactions_new %>% filter(!id %in% rmv_df$id)
# 
# rm(rmv_df)
```

## Save new interactions data frame

```{r}
saveRDS(object = interactions_new, file = paste0(out_dir,"/interactions_all_", date,".rds"))
```

## Tests

### Test: Suspicious regions with a higher number of interacting cell types than active cell types.

CONCLUSION: THESE SUSPICIOUS REGIONS ARE CAUSED BY THE MERGING PROCEDURE. THEY ONLY AFFECT THE REGION NOT CTS NOR HK. FOR NOW WE CAN FILTER THEM OUT.

```{r}
tt <- interactions_new %>% filter(n_ct_interactions > 2 & test_n_celltypes == 2) 
```

there are some regions that overlaps with all the 50 cell types. Let's verify this

```{r}
sel_region <- "chr20:17680712-17681155_RRBP1"
tmp_region <- interactions_new %>% filter(id == sel_region) %>% as.data.frame()

tmp_idx <- WhichRegionsOverlap(query_enhancer = tmp_region,
                               subject_enhancer = regions_all)
tmp_overlaps <- regions_all[subjectHits(tmp_idx),]
tmp_overlaps <- tmp_overlaps %>% filter(TargetGene %in% tmp_region$gene)
length(unique(tmp_overlaps$CellType))
```

This regions in my overlapping method was found in 2 cell types, but in this analysis it overlaps with all of them... 
I think it caused by the overlapping procedure where regions dissapear in their sorroundings and then they appear as only in 2 cell types. In such case this region sould be close to a HK regions. lets verify

```{r}
tmp_regions <- interactions_new %>% filter(chr == "chr20" & start >= 17600000)
```

Yes!! this regions is close to a region which is found in 38 cell types. And the next one is found in 50 cell type

Maybe it's better to filter out these regions?

Let's try to get a cell type specific one 

```{r}
tt <- interactions_new %>% filter(test_n_celltypes == 1) 
```

all of them are OK. I think the problem is the noisy ones which are not cell type-specific nor HK regions.

Let's see how many we would lose if we filter them

```{r}
interactions_new %>% filter(!duplicated(region_id)) %>% ggplot(mapping = aes(x = test_n_celltypes)) + geom_histogram(stat = "count")
interactions_new %>% filter(!duplicated(region_id)) %>% pull(test_n_celltypes) %>% table()
```

```{r}
#get the ids of regions that contain more number of interacting cell types than number of active cell types
tmp_regions <- interactions_new %>%
  filter(n_ct_interactions > n_celltypes) %>%
  filter(!duplicated(region_id))
#filter out suspicious regions 
tt <- interactions_new %>% filter(!region_id %in% tmp_regions$region_id) %>%   filter(!duplicated(region_id)) 

tt %>% ggplot(mapping = aes(x = test_n_celltypes)) + geom_histogram(stat = "count")
tt %>% pull(test_n_celltypes) %>% table()
```

I think they look good!!

I will filter them for further analysis.

### Test 1: Suspicious interactions

CONCLUSIONS: THE SUSPICIOUS INTERACTIONS (NOT APPEARED BEFORE) ARE CAUSED BY THE MERGING METHOD. WE CAN KEEP THEM FOR NOW.

It looks like there are interactions that I didn't have before... I am not sure why...

```{r}
#short test
any(duplicated(n_ct_interaction$id))
any(duplicated(interactions_cp$id))

sum(interactions_cp$id %in% n_ct_interaction$id) == nrow(interactions_cp)
```

```{r}
suspicious_ids <- which(!n_ct_interaction$id %in% interactions_cp$id)
suspicious_ids <- n_ct_interaction[suspicious_ids,]
```

```{r}
interactions_cp %>% filter(id == "chr1:1013217-1013716_ISG15")
interactions_cp %>% filter(region_id == "chr1:1013217-1013716")
```
```{r}
test_interactions <- interactions_cp %>% filter(id == "chr1:1013217-1013716_ISG15")
#get overlapping interactions in the original data set
ovl_idx <- WhichRegionsOverlap(query_enhancer = test_interactions,
                                 subject_enhancer = regions_all)
#get overlap regions from the original raw data
all_overlaps <- regions_all[subjectHits(ovl_idx),] %>% filter(TargetGene == test_interactions$gene) 
length(unique(all_overlaps$CellType))
```

```{r}
all_overlaps %>% count(TargetGene, CellType) 
```

How to improve the count ...

```{r}
test_interactions <- interactions_cp %>% filter(id == n_ct_interaction$id[1])
#get overlapping interactions in the original data set
ovl_idx <- WhichRegionsOverlap(query_enhancer = test_interactions,
                                 subject_enhancer = regions_all)
#get overlap regions from the original raw data
all_overlaps <- regions_all[subjectHits(ovl_idx),]
all_overlaps %>% group_by(TargetGene) %>% count(CellType) %>% ungroup() %>% count(TargetGene)
```

```{r}
all_overlaps %>% filter(TargetGene == "GPKOW") %>% pull(CellType) %>% unique()
```

### Test 3.

Test unique cell types..

```{r}
sum(unique(regions_all$CellType) %in% interactions_cp$celltypes)
unique(regions_all$CellType)
```

