---
title: "P-P Interactions, analysis by graph"
author: "Martin Loza"
date: "2023/09/18
---

For the set of core-promoters, I would like to make a graph analysis of thein interactions. I would like to use the n_ct_interactions as a weight for the edges.


NOTES:


```{r setup}
library(dplyr)
library(ggplot2)
library(patchwork)
library(stringr)
library(GenomicRanges)
library(igraph)
# library(org.Hs.eg.db)
# library(AnnotationDbi)
library(ggraph)

## Global variables
in_dir  <- "~/Documents/Projects/Enhancer_Grammar/Analysis/Weekly_analyses/2023_09/2023_09_18/Results/"
out_dir <- "~/Documents/Projects/Enhancer_Grammar/Analysis/Weekly_analyses/2023_09/2023_09_18/Results/"

date = 2309018
red = "#CC6677"
blue = "#6699CC"
yellow = "#DDCC77"

#Functions
source(file = "~/Documents/Projects/Enhancer_Grammar/Analysis/Functions.R")
source(file = "~/Documents/Projects/Enhancer_Grammar/Analysis/Functions_Visualizations.r")

#Local functions 
GetGenesAnnotations <- function(keys = NULL, keys_type = "SYMBOL", annotations = c("SYMBOL","ENSEMBL","ENTREZID"), organism = "human"){
  if(organism == "human"){
    gene_annotations <- AnnotationDbi::select(x = org.Hs.eg.db, 
                                              keys = keys,
                                              columns = annotations,
                                              keytype = keys_type)
  }
  return(gene_annotations)
}

GetGenerationEdges <- function(edges = NULL, init_nodes = NULL, n_generations = 2){
  
  #loop. In each generation get the edges associated to init nodes
  for(i in 2:n_generations){
    tmp_df <- edges %>% filter(gene_from %in% all_of(init_nodes) | gene_to %in% all_of(init_nodes))
    init_nodes <- union(tmp_df$gene_from, tmp_df$gene_to)
  }

  return(tmp_df)
}
```

## Load data

Load annotated interactions 

```{r}
interactions <- readRDS(paste0(in_dir,"/interactions_all_230912.Rds"))
interactions %>% filter(!duplicated(region_id)) %>% pull(n_celltypes) %>% table()
```

## Set up

I would like to select interactions only for CP 

```{r}
interactions <- interactions %>% filter(ann_v1 == "CP")
interactions$ann_v1 <- droplevels(interactions$ann_v1)
interactions %>% filter(!duplicated(region_id)) %>% pull(n_celltypes) %>% table()
```

Check for duplicated interactions 

```{r}
any(duplicated(interactions$id))
```

## Analysis

### Directed graph

For now, let's assume that the regions corresponds to an annotated gene
Later, let's check for divergent genes too. For now, let's use the nearest genes

We need to get a data frame with the edges and the weight information

```{r}
#set new genes with direction
interactions <- interactions %>% 
  mutate(gene_from = nearest_gene, gene_to = gene) %>%
  mutate(weight = n_ct_interactions/50) #normalized weights
#get the edges data frames
edges_df <- interactions %>% dplyr::select(gene_from, gene_to, weight)
head(edges_df)
```

Get the directed graph

```{r}
#get graph from edges
graph_dir<- graph.data.frame(d = edges_df, directed = TRUE)
```

Short tests

```{r}
is_weighted(graph_dir)
tt <- E(graph_dir)$weight
summary(tt)
rm(tt)
```

Let's analyse the adjacency matrix

```{r}
#get adjacency matrix
adj_matrix <- get.adjacency(graph_dir, attr = "weight" )# %>% as.matrix()
adj_matrix[1:10, 1:10]
```
We can see how the weight has affected the interactions. 

It looks like in the case of weighted graphs, the degree it's called strength. 
Let's get the strength of the nodes

```{r}
strength <- strength(graph = graph_dir, mode = "all", loops = TRUE)
strength <- data.frame(gene = names(strength), strength = strength)
```

Let's try to select the top nodes

```{r}
top_nodes <- strength %>% slice_max(order_by = strength, n = 100) %>% arrange(desc(strength))
top_nodes$gene
```
#### Plot top nodes

For the top nodes, let's get the nodes until the second generation

```{r}
sel_edges <- GetGenerationEdges(edges = edges_df, init_nodes = "WDR74", n_generations = 3)
```

```{r}
sel_graph <- graph.data.frame(d = sel_edges, directed = TRUE)
is_weighted(sel_graph)

sel_strength <- strength(graph = sel_graph, mode = "all", loops = TRUE)
sel_strength <- data.frame(gene = names(sel_strength), strength = sel_strength)

#summary strength
summ <- summary(sel_strength$strength)
summ

#select top nodes
sel_nodes <- sel_strength %>% slice_max(order_by = strength, n = 5)
# sel_nodes <- sel_strength %>% filter(strength >= summ["3rd Qu."])
sel_nodes$gene
```
Let's plot

```{r fig.asp=1}
sel_graph %>%
  ggraph(layout = "centrality",
         # centrality = evcent(sel_graph)$vector) +
         centrality = hub.score(sel_graph)$vector) +
  geom_edge_link0(aes(edge_alpha = E(sel_graph)$weight), edge_colour = "gray70") +
  geom_node_point(aes(size = sel_strength$strength), show.legend = FALSE,  color = blue) +
  theme_void() + 
  geom_node_text(aes(label = name, filter = name %in% sel_nodes$gene ), 
                 color = 'black', 
                 size = 3)+ theme(legend.position = "none")
```

### HK PPI graph

For now, let's assume that the regions corresponds to an annotated gene
Later, let's check for divergent genes too. For now, let's use the nearest genes

We need to get a data frame with the edges and the weight information

```{r}
#set new genes with direction
interactions <- interactions %>% filter(n_ct_interactions > 45) %>% 
  mutate(gene_from = nearest_gene, gene_to = gene) %>%
  mutate(weight = n_ct_interactions/50) #normalized weights
#get the edges data frames
edges_df <- interactions %>% dplyr::select(gene_from, gene_to, weight)
head(edges_df)
```

Get the directed graph

```{r}
#get graph from edges
graph_dir<- graph.data.frame(d = edges_df, directed = TRUE)
```

Short tests

```{r}
is_weighted(graph_dir)
tt <- E(graph_dir)$weight
summary(tt)
rm(tt)
```

Let's analyse the adjacency matrix

```{r}
#get adjacency matrix
adj_matrix <- get.adjacency(graph_dir, attr = "weight" )# %>% as.matrix()
adj_matrix[1:10, 1:10]
```
We can see how the weight has affected the interactions. 

It looks like in the case of weighted graphs, the degree it's called strength. 
Let's get the strength of the nodes

```{r}
strength <- strength(graph = graph_dir, mode = "all", loops = TRUE)
strength <- data.frame(gene = names(strength), strength = strength)
```

Let's try to select the top nodes

```{r}
top_nodes <- strength %>% slice_max(order_by = strength, n = 100) %>% arrange(desc(strength))
top_nodes$gene
```
#### Plot top nodes

For the top nodes, let's get the nodes until the second generation

```{r}
sel_edges <- GetGenerationEdges(edges = edges_df, init_nodes = "USP5", n_generations = 3)
```

```{r}
sel_graph <- graph.data.frame(d = sel_edges, directed = TRUE)
is_weighted(sel_graph)

sel_strength <- strength(graph = sel_graph, mode = "all", loops = TRUE)
sel_strength <- data.frame(gene = names(sel_strength), strength = sel_strength)

#summary strength
summ <- summary(sel_strength$strength)
summ

#select top nodes
sel_nodes <- sel_strength %>% slice_max(order_by = strength, n = 5)
# sel_nodes <- sel_strength %>% filter(strength >= summ["3rd Qu."])
sel_nodes$gene
```
Let's plot

```{r fig.asp=1}
sel_graph %>%
  ggraph(layout = "centrality",
         # centrality = evcent(sel_graph)$vector) +
         centrality = hub.score(sel_graph)$vector) +
  geom_edge_link0(aes(edge_alpha = E(sel_graph)$weight), edge_colour = "gray70") +
  geom_node_point(aes(size = sel_strength$strength), show.legend = FALSE,  color = blue) +
  theme_void() + 
  geom_node_text(aes(label = name, filter = name %in% sel_nodes$gene ), 
                 color = 'black', 
                 size = 3)+ theme(legend.position = "none")
```

```{r}
names(sel_nodes)

```



```{r}

sel_graph %>%
  ggraph(layout = "centrality", centrality = closeness(sel_graph)) +
  # draw_circle(use = "cent") +
  # ggraph(layout = "kk") +
  geom_edge_link0(aes(edge_alpha = E(sel_graph)$weight), edge_colour = "gray") +
  geom_node_point(aes(size = sel_strength), show.legend = TRUE,  color = "gray") +
  theme_void()
```



```{r}
thr <- summary(sel_strength$strength)["3rd Qu."]
ggraph(graph = sel_graph) + 
    geom_edge_link() + 
    # geom_node_point(aes(size = sel_strength$strength), color = "gray") #+ 
    theme_graph( fg_text_colour = 'white') #+ 
  # geom_node_text(aes(label = name, filter = strength > thr), 
  #                color = 'black', 
  #                size = 3)
```







### Plot

```{r}
#get graph from edges
graph_directed <- graph.data.frame(d = edges_df, directed = TRUE)
```


Let's try to select the top nodes

```{r}
degree_out <- igraph::degree(graph = graph_directed, mode = "all", loops = TRUE, normalized = TRUE)
degree_out <- data.frame(gene = names(degree_out), degree = degree_out)
top_degree <- degree_out %>% slice_max(order_by = degree, n = 3)
```

```{r}
test_df <- edges_df %>% filter(gene_from %in% top_degree$gene[1] | gene_to %in% top_degree$gene[1])
test_graph <- graph.data.frame(d = test_df, directed = TRUE)
```

```{r}
degree <- igraph::degree(graph = test_graph, mode = "all", loops = TRUE, normalized = TRUE)
d_sum <- summary(degree)
```

```{r}
thr <- d_sum["3rd Qu."]
ggraph(test_graph, layout = "centrality",cent=degree) + 
    geom_edge_link0( edge_alpha = 0.1) + 
    geom_node_point(aes(size = degree), color = "gray") + 
    theme_graph( fg_text_colour = 'white') + 
  geom_node_text(aes(label = name, filter = degree > thr), 
                 color = 'black', 
                 size = 3)
```

```{r}
test_df <- edges_df %>% filter(gene_from %in% top_degree$gene[2] | gene_to %in% top_degree$gene[2])
test_graph <- graph.data.frame(d = test_df, directed = TRUE)
```

```{r}
degree <- igraph::degree(graph = test_graph, mode = "all", loops = TRUE, normalized = TRUE)
d_sum <- summary(degree)
```

```{r}
thr <- d_sum["3rd Qu."]
ggraph(test_graph, layout = "centrality",cent=degree) + 
    geom_edge_link0(edge_alpha = 0.1) + 
    geom_node_point(aes(size = degree), color = "gray") + 
    theme_graph( fg_text_colour = 'white') + 
  geom_node_text(aes(label = name, filter = degree > thr), 
                 color = 'black', 
                 size = 3)
```

```{r}
test_df <- edges_df %>% filter(gene_from %in% top_degree$gene[3] | gene_to %in% top_degree$gene[3])
test_graph <- graph.data.frame(d = test_df, directed = TRUE)
```

```{r}
degree <- igraph::degree(graph = test_graph, mode = "all", loops = TRUE, normalized = TRUE)
d_sum <- summary(degree)
```

```{r}
thr <- d_sum["3rd Qu."]
ggraph(test_graph, layout = "centrality",cent=degree) + 
    geom_edge_link0( edge_alpha = 0.1) + 
    geom_node_point(aes(size = degree), color = "gray") + 
    theme_graph( fg_text_colour = 'white') + 
  geom_node_text(aes(label = name, filter = degree > thr), 
                 color = 'black', 
                 size = 3)
```

```{r}
test_df <- edges_df %>% filter(gene_from %in% top_degree$gene[1:3] | gene_to %in% top_degree$gene[1:3])
test_graph <- graph.data.frame(d = test_df, directed = TRUE)
```

```{r}
degree <- igraph::degree(graph = test_graph, mode = "all", loops = TRUE, normalized = TRUE)
d_sum <- summary(degree)
```

```{r}
thr <- d_sum["3rd Qu."]
ggraph(test_graph, layout = 'kk') + 
    geom_edge_link0( edge_alpha = 0.1) + 
    geom_node_point(aes(size = degree), color = "gray") + 
    theme_graph( fg_text_colour = 'white') + 
  geom_node_text(aes(label = name, filter = degree > thr), 
                 color = 'black', 
                 size = 3)
```

## Test

### Testing

Testing the in/out edges from the symmetric adjacency matrix of genes

```{r}
#this means tha GNB1 target itself 4 times. Let's verify this
tt <- interactions %>% filter(gene_from == "GNB1" & gene_to == "CFAP74")
ttt <- interactions %>% filter(gene_to == "GNB1" & gene_from == "CFAP74")
```



Let's try using directed graphs
```{r}
#get graph from edges
graph_interactions <- graph.data.frame(d = edges_df, directed = TRUE)
#get adjacency matrix
sym_matrix <- get.adjacency(graph_interactions)# %>% as.matrix()
sym_matrix[1:6, 1:6]
```

Let's count the number of interactions for each gene. We need to set as zero the upper triangle

```{r}
sym_matrix_zeros <- sym_matrix
sym_matrix_zeros[upper.tri(sym_matrix_zeros)] <- 0
sym_matrix_zeros[1:6, 1:6]
```
```{r}
total_interactions_row <- apply(X = sym_matrix_zeros, MARGIN = 1, FUN = sum)
total_interactions_row[1:6]
```

Let's count the number of interactions for each gene. We need to set as zero the upper triangle

```{r}
sym_matrix_zeros <- sym_matrix
sym_matrix_zeros[lower.tri(sym_matrix_zeros)] <- 0
sym_matrix_zeros[1:6, 1:6]
```

```{r}
total_interactions_col <- apply(X = sym_matrix_zeros, MARGIN = 2, FUN = sum)
total_interactions_col[1:6]
```

```{r}
total_interactions <- total_interactions_col + total_interactions_row
```

```{r}
total_interactions %>% sort(decreasing = TRUE)
```

```{r}
plot(x = graph_interactions)
```
```{r}
library(ggnetwork)
n <- ggnetwork(x = graph_interactions
```

```{r}
head(n)
```


```{r fig.width=10}
ggplot(n, aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_edges(color = "black") +
  # geom_nodelabel(aes(label = LETTERS[ name ]),
  #                fontface = "bold") +
  theme_blank()
```







## Tests

#### Test using indirected graph

It didn't work as I expected... 
The final count doesn't get what I want

```{r}
#get graph from edges
graph_interactions <- graph.data.frame(d = edges_df, directed = FALSE)
#get adjacency matrix
sym_matrix <- get.adjacency(graph_interactions) %>% as.matrix()
sym_matrix[1:6, 1:6]
```

Let's count the number of interactions for each gene. We need to set as zero the upper triangle

```{r}
sym_matrix_zeros <- sym_matrix
sym_matrix_zeros[upper.tri(sym_matrix_zeros)] <- 0
sym_matrix_zeros[1:6, 1:6]
```

```{r}
total_interactions <- apply(X = sym_matrix_zeros, MARGIN = 1, FUN = sum)
total_interactions[1:6]
```
```{r}
# Create a sample matrix
mat <- matrix(1:9, nrow = 3)

# Convert the upper triangle to zeros
mat[upper.tri(mat)] <- 0

# Print the modified matrix
print(mat)
```



## Tests

### Test for adjacency matrix

let's use the igraph package.

```{r}
# Create a simple undirected graph
edges <- data.frame(
  from = c("1", "2", "3", "4", "4"),
  to = c("2", "4", "1", "2", "2")
)

as.matrix(edges)

# Create a graph object
graph_directed <- graph.data.frame(edges, directed = TRUE)
graph_indirected <- graph.data.frame(edges, directed = FALSE)

# Get the adjacency matrices
sym_matrix <- get.adjacency(graph_indirected)
asym_matrix <- get.adjacency(graph_directed)

# sym_matrix
asym_matrix
sym_matrix
# as.matrix(sym_matrix) & as.matrix(asym_matrix)
```

I think I can use a symetric matrix from an indirected graph. Then count those interactions with more than one value 














adjacency matrix
```{r}
#get the universe of genes
genes_universe = union(interactions$gene_enhancer, y = interactions$gene_target)
#short test
any(duplicated(genes_universe))


```

```{r}
library(igraph)
```











For each interaction

```{r}
#init number of celltypes of interactions
n_ct_interaction <- matrix(data = 0, nrow = nrow(interactions), ncol = 1)
#get cell types as a matrix to ease their index
n_celltypes_all <- matrix(regions_all$CellType, ncol = 1)

#for each interaction
# for(i in seq_len(nrow(interactions))){
  for(i in 158426:nrow(interactions)){
  #for tests
  # for(i in 1:10){
  # i = 1
  
  #get current interaction
  c_interaction = interactions[i,]
  #get overlapping interactions in the original data set
  ovl_idx <- WhichRegionsOverlap(query_enhancer = c_interaction,
                                 subject_enhancer = regions_all)
  #get the current cell types
  c_ct <- n_celltypes_all[subjectHits(ovl_idx),1]
  #count the current cell types
  n_ct <- length(unique(c_ct))
  #assign the number of cell types
  n_ct_interaction[i,1] <- n_ct
  
}

#rmv unnecesary objects
rm( c_interaction, ovl_idx, c_ct, n_ct)
```




















How many CREs target each genes. I would like to know by cell types

```{r}
#count number of interactions
interactions <- interactions %>% 
  group_by(gene) %>%
  count() %>%
  right_join(.,y = interactions, by = "gene")
#rename for downstream analyses
interactions <- interactions %>% 
  dplyr::rename(n_interactions = n)
```

Now I would like to know how many HK regions target each cp

```{r}
tmp_interactions <- interactions %>% 
  group_by(gene) %>% 
  count(type == "HK")
#add colnames for further pipelines
colnames(tmp_interactions) <- c("gene", "is_HKG", "n_HK_CREs")
#select HK-interactions 
tmp_interactions <-tmp_interactions %>% filter(is_HKG == TRUE) %>% select(gene, n_HK_CREs)
#transfer information to main data frame
interactions <- interactions %>% left_join(y = tmp_interactions, by = "gene")
#Add zeros to NAs
idx <- which(is.na(interactions$n_HK_CREs))
interactions$n_HK_CREs[idx] <- 0

rm(tmp_interactions, idx)
```

Now, I would like to obtain a percentage of number of HK involved on each cp_gene

```{r}
interactions <- interactions %>% mutate(per_HK_CREs = n_HK_CREs/n_interactions)
```

Now I would like to know how many CTS regions target each cp

```{r}
tmp_interactions <- interactions %>% 
  group_by(gene) %>% 
  count(n_celltypes == 1)
#add colnames for further pipelines
colnames(tmp_interactions) <- c("gene", "is_CTS", "n_CTS_CREs")
#select CTS-interactions 
tmp_interactions <-tmp_interactions %>% filter(is_CTS == TRUE) %>% select(gene, n_CTS_CREs)
#transfer information to main data frame
interactions <- interactions %>% left_join(y = tmp_interactions, by = "gene")
#Add zeros to NAs
idx <- which(is.na(interactions$n_CTS_CREs))
interactions$n_CTS_CREs[idx] <- 0

rm(tmp_interactions, idx)
```

Now, I would like to obtain a percentage of number of CTS involved on each cp_gene

```{r}
interactions <- interactions %>% mutate(per_CTS_CREs = n_CTS_CREs/n_interactions)
```

I would like to annotate which gene is a HKG for further comparisons 

```{r}
interactions <- interactions %>% mutate(is_HKG = ifelse(test = gene %in% hkg$gene_name, yes = TRUE, no = FALSE))
```

I would like to annote which gene is found to have a housekeeping core promoters

```{r}
interactions <- interactions %>%
  mutate(is_HK_CP = ifelse(test = gene %in% hk_cp_ann$gene_name, yes = TRUE, no = FALSE))
```

I would like to annotate how many CP or P and how many ESD, ELD, or ESLD target each gene

```{r}
interactions <- interactions %>% 
  group_by(gene) %>%
  summarise(n_P = sum(ann_v1 == "CP" | ann_v1 == "P"),
            n_E = sum(ann_v1 == "ESD" | ann_v1 == "ELD"| ann_v1 == "ESLD")) %>%
  right_join(y = interactions, by = "gene")

#proportion of enhancers and promoters
interactions <- interactions %>% mutate(p_E = n_E/n_interactions)
interactions <- interactions %>% mutate(p_P = n_P/n_interactions)
```

All together

```{r}
#create new label
interactions <- interactions %>% mutate(type_target = "Other")
idx <- which(interactions$is_HK_CP)
interactions$type_target[idx] <- "nonHKG"
idx <- which(interactions$is_HKG)
interactions$type_target[idx] <- "HKG"

interactions <- interactions %>% 
  mutate(type_target = factor(type_target, 
                              levels = c("Other", "HKG", "nonHKG") ))
table(interactions$type_target)
```

```{r}
saveRDS(object = interactions, file = paste0(out_dir,"/interactions_",date,".rds"))
```


### Plots

```{r}
plot_df <- interactions %>% filter(!duplicated(gene))
plot_df$test_n_celltypes <- factor(plot_df$test_n_celltypes, levels = 1:50)
plot_df$n_celltypes <- factor(plot_df$n_celltypes, levels = c(1:44,50))
table(plot_df$type_target)
```
### Number of interactions

```{r}
my_comparison <- list(c("Other", "HKG"), 
                      c("Other", "nonHKG"),
                      c("HKG", "nonHKG"))

p <- plot_df %>% 
  MyComparisonPlot(wt = "n_interactions",
                   compare_wt = "type_target", 
                   palette =  c("gray", red, blue), 
                   adjust = 1,
                   comparisons = my_comparison,
                   label = "p.signif",
                   vjust = 0.5,
                   box_width = 0.15) + 
  ylab(expression("Number of interactions")) + 
  NoLegend() +
  theme(axis.title.x = element_blank(), plot.title = element_blank()) +
  theme(text = element_text(size = text_size)) +
  geom_jitter(size = 0.1, color = "black", alpha = 0.05)

p

# ggsave(filename = paste0("n_interactions_", date, ".png"), plot = p, device = "png",
#        path = out_dir, width = (fig_size), height = fig_size, dpi = dpi )
# ggsave(filename = paste0("n_interactions_", date, ".pdf"), plot = p, device = "pdf",
#        path = out_dir, width = (fig_size), height = fig_size, dpi = dpi )

```

```{r}
#statistics other
plot_df %>% filter(type_target == "Other") %>% pull(n_interactions) %>% summary()
#statistics other
plot_df %>% filter(type_target == "HKG") %>% pull(n_interactions) %>% summary()
#statistics other
plot_df %>% filter(type_target == "nonHKG") %>% pull(n_interactions) %>% summary()
```
Genes with ubiquitously active promoter are predicted as to have more interactions

### Distance of interactions

```{r fig.width=5, fig.height=5}
my_comparison <- list(c("Other", "HKG"), 
                      c("Other", "nonHKG"),
                      c("HKG", "nonHKG"))

p <- plot_df %>% 
  MyComparisonPlot(wt = "distance",
                   compare_wt = "type_target", 
                   palette =  c("gray", red, blue), 
                   adjust = 3,
                   comparisons = my_comparison,
                   label = "p.signif",
                   vjust = 0.5,
                   box_width = 0.15) + 
  scale_y_continuous(trans = "log10") +
  ylab(expression("Distance")) + 
  NoLegend() +
  theme(axis.title.x = element_blank(), plot.title = element_blank()) +
  theme(text = element_text(size = text_size)) +
  geom_jitter(size = 0.1, color = "black", alpha = 0.05)

p

# ggsave(filename = paste0("n_interactions_", date, ".png"), plot = p, device = "png",
#        path = out_dir, width = (fig_size), height = fig_size, dpi = dpi )
# ggsave(filename = paste0("n_interactions_", date, ".pdf"), plot = p, device = "pdf",
#        path = out_dir, width = (fig_size), height = fig_size, dpi = dpi )

```

```{r}
#statistics other
plot_df %>% filter(type_target == "Other") %>% pull(distance) %>% summary()
#statistics other
plot_df %>% filter(type_target == "HKG") %>% pull(distance) %>% summary()
#statistics other
plot_df %>% filter(type_target == "nonHKG") %>% pull(distance) %>% summary()
```
Distance between cts-CREs and HK-CREs

```{r}
plot_df %>% 
  filter(n_celltypes == 1 | n_celltypes == 50) %>% 
  ggplot(mapping = aes(x = distance, fill = n_celltypes)) +
  geom_histogram(position = "dodge") + 
  scale_x_continuous(trans = "pseudo_log")
```


### Number and proportion of promoters

I would like to analyze this from the perspective of number of enhancers an number of promoters

```{r}
my_comparison <- list(c("Other", "HKG"), 
                      c("Other", "nonHKG"),
                      c("HKG", "nonHKG"))

p1 <- plot_df %>% 
  MyComparisonPlot(wt = "p_P",
                   compare_wt = "type_target", 
                   palette =  c("gray", red, blue), 
                   adjust = 1,
                   comparisons = my_comparison,
                   label = "p.signif",
                   vjust = 0.3,
                   box_width = 0.15) + 
  ylab(expression("Proportion of promoters")) + 
  NoLegend() +
  theme(axis.title.x = element_blank(), plot.title = element_blank()) +
  theme(text = element_text(size = text_size)) +
  geom_jitter(size = 0.1, color = "black", alpha = 0.05)

p2 <- plot_df %>% 
  MyComparisonPlot(wt = "n_P",
                   compare_wt = "type_target", 
                   palette =  c("gray", red, blue), 
                   adjust = 1,
                   comparisons = my_comparison,
                   label = "p.signif",
                   vjust = 0.3,
                   box_width = 0.15) + 
  ylab(expression("Number of promoters")) + 
  NoLegend() +
  theme(axis.title.x = element_blank(), plot.title = element_blank()) +
  theme(text = element_text(size = text_size)) +
  geom_jitter(size = 0.1, color = "black", alpha = 0.05)

p1 + p2

# ggsave(filename = paste0("n_interactions_", date, ".png"), plot = p, device = "png",
#        path = out_dir, width = (fig_size), height = fig_size, dpi = dpi )
# ggsave(filename = paste0("n_interactions_", date, ".pdf"), plot = p, device = "pdf",
#        path = out_dir, width = (fig_size), height = fig_size, dpi = dpi )

```

```{r}
my_comparison <- list(c("HKG", "nonHKG"))

p1 <- plot_df %>% 
  filter(type_target != "Other") %>% 
  MyComparisonPlot(wt = "p_P",
                   compare_wt = "type_target", 
                   palette =  c("gray", red, blue), 
                   adjust = 1,
                   comparisons = my_comparison,
                   label = "p.signif",
                   vjust = 0.3,
                   box_width = 0.15) + 
  ylab(expression("Proportion of promoters")) + 
  NoLegend() +
  theme(axis.title.x = element_blank(), plot.title = element_blank()) +
  theme(text = element_text(size = text_size)) +
  geom_jitter(size = 0.1, color = "black", alpha = 0.05)

p2 <- plot_df %>% 
  filter(type_target != "Other") %>% 
  MyComparisonPlot(wt = "n_P",
                   compare_wt = "type_target", 
                   palette =  c("gray", red, blue), 
                   adjust = 1,
                   comparisons = my_comparison,
                   label = "p.signif",
                   vjust = 0.3,
                   box_width = 0.15) + 
  ylab(expression("Number of promoters")) + 
  NoLegend() +
  theme(axis.title.x = element_blank(), plot.title = element_blank()) +
  theme(text = element_text(size = text_size)) +
  geom_jitter(size = 0.1, color = "black", alpha = 0.05)

p1 + p2

```

```{r}
#statistics other
plot_df %>% filter(type_target == "Other") %>% pull(p_P) %>% summary()
#statistics other
plot_df %>% filter(type_target == "HKG") %>% pull(p_P) %>% summary()
#statistics other
plot_df %>% filter(type_target == "nonHKG") %>% pull(p_P) %>% summary()
```

```{r}
#statistics other
plot_df %>% filter(type_target == "HKG") %>% pull(n_P) %>% summary()
#statistics other
plot_df %>% filter(type_target == "nonHKG") %>% pull(n_P) %>% summary()
```

### Number and proportion of Enhancers

```{r}
my_comparison <- list(c("HKG", "nonHKG"))

p1 <- plot_df %>% 
  MyComparisonPlot(wt = "p_E",
                   compare_wt = "type_target", 
                   palette =  c("gray", red, blue), 
                   adjust = 1,
                   comparisons = my_comparison,
                   label = "p.signif",
                   vjust = 0.3,
                   box_width = 0.15) + 
  ylab(expression("Proportion of enhancers")) + 
  NoLegend() +
  theme(axis.title.x = element_blank(), plot.title = element_blank()) +
  theme(text = element_text(size = text_size)) +
  geom_jitter(size = 0.1, color = "black", alpha = 0.05)

p2 <- plot_df %>% 
  MyComparisonPlot(wt = "n_E",
                   compare_wt = "type_target", 
                   palette =  c("gray", red, blue), 
                   adjust = 1,
                   comparisons = my_comparison,
                   label = "p.signif",
                   vjust = 0.3,
                   box_width = 0.15) + 
  ylab(expression("Number of enhancers")) + 
  NoLegend() +
  theme(axis.title.x = element_blank(), plot.title = element_blank()) +
  theme(text = element_text(size = text_size)) +
  geom_jitter(size = 0.1, color = "black", alpha = 0.05)

p1 + p2

```

```{r}
my_comparison <- list(c("HKG", "nonHKG"))

p1 <- plot_df %>% 
  filter(type_target != "Other") %>% 
  MyComparisonPlot(wt = "p_E",
                   compare_wt = "type_target", 
                   palette =  c("gray", red, blue), 
                   adjust = 1,
                   comparisons = my_comparison,
                   label = "p.signif",
                   vjust = 0.3,
                   box_width = 0.15) + 
  ylab(expression("Proportion of enhancers")) + 
  NoLegend() +
  theme(axis.title.x = element_blank(), plot.title = element_blank()) +
  theme(text = element_text(size = text_size)) +
  geom_jitter(size = 0.1, color = "black", alpha = 0.05)

p2 <- plot_df %>% 
  filter(type_target != "Other") %>% 
  MyComparisonPlot(wt = "n_E",
                   compare_wt = "type_target", 
                   palette =  c("gray", red, blue), 
                   adjust = 1,
                   comparisons = my_comparison,
                   label = "p.signif",
                   vjust = 0.3,
                   box_width = 0.15) + 
  ylab(expression("Number of enhancers")) + 
  NoLegend() +
  theme(axis.title.x = element_blank(), plot.title = element_blank()) +
  theme(text = element_text(size = text_size)) +
  geom_jitter(size = 0.1, color = "black", alpha = 0.05)

p1 + p2

```

```{r}
#statistics other
plot_df %>% filter(type_target == "Other") %>% pull(p_E) %>% summary()
#statistics other
plot_df %>% filter(type_target == "HKG") %>% pull(p_E) %>% summary()
#statistics other
plot_df %>% filter(type_target == "nonHKG") %>% pull(p_E) %>% summary()
```

```{r}
#statistics other
plot_df %>% filter(type_target == "HKG") %>% pull(n_E) %>% summary()
#statistics other
plot_df %>% filter(type_target == "nonHKG") %>% pull(n_E) %>% summary()
```

### Number of enhancers and number of promoters 

```{r}
#housekeeping core promoters
plot_df %>% filter(is_HK_CP) %>%
  ggplot(mapping = aes(y = n_P, x = n_E)) +
  geom_point(alpha = 0.1)+ coord_fixed(xlim = c(0,60), ylim = c(0,60)) + theme_light() + ggtitle("HK-CPs")
#non-housekeeping core promoters
plot_df %>% filter(!is_HK_CP) %>%
  ggplot(mapping = aes(y = n_P, x = n_E)) +
  geom_point(alpha = 0.1) + coord_fixed(xlim = c(0,60), ylim = c(0,60))
#housekeeping genes
plot_df %>% filter(is_HKG) %>%
  ggplot(mapping = aes(y = n_P, x = n_E)) +
  geom_point(alpha = 0.1) + coord_fixed(xlim = c(0,60), ylim = c(0,60))
#housekeeping core promoters, except housekeeping genes
plot_df %>% filter(is_HK_CP) %>%
  filter(!is_HKG) %>%
  ggplot(mapping = aes(y = n_P, x = n_E)) +
  geom_point(alpha = 0.1) + coord_fixed(xlim = c(0,60), ylim = c(0,60))
```

```{r}
tmp_df <- plot_df %>% mutate(combination = paste(n_P, n_E, sep = "_"))
frequency_others <- tmp_df %>% group_by(combination) %>% count() %>% arrange(desc(n)) %>% rename(counts = n )
most_frequent <- frequency_others %>% ungroup() %>% arrange(desc(counts)) %>% slice_max(counts)

tmp_df %>% 
  left_join(., y = frequency_others, by = "combination") %>%
  ggplot(mapping = aes(x = n_P, y = n_E, color = counts)) +
  geom_point() +
  coord_fixed(xlim = c(0,60), ylim = c(0,60)) + theme_light() + ggtitle("All genes", subtitle = most_frequent$combination) + scale_color_gradient2( low= blue, mid = "gray", high=red)
```

```{r}
tmp_df <- plot_df %>% filter(is_HK_CP) %>% mutate(combination = paste(n_P, n_E, sep = "_"))
frequency_others <- tmp_df %>% group_by(combination) %>% count() %>% arrange(desc(n)) %>% rename(counts = n )
most_frequent <- frequency_others %>% ungroup() %>% arrange(desc(counts)) %>% slice_max(counts)

tmp_df %>% 
  left_join(., y = frequency_others, by = "combination") %>%
  ggplot(mapping = aes(x = n_P, y = n_E, color = counts)) +
  geom_point() +
  coord_fixed(xlim = c(0,60), ylim = c(0,60)) + theme_light() + ggtitle("HK-CPs", subtitle = most_frequent$combination) + scale_color_gradient2( low= blue, mid = "gray", high=red)
```

```{r}
#non-housekeeping core promoters
tmp_df <- plot_df %>% filter(!is_HK_CP) %>% mutate(combination = paste(n_P, n_E, sep = "_"))
frequency_others <- tmp_df %>% group_by(combination) %>% count() %>% arrange(desc(n)) %>% rename(counts = n )
most_frequent <- frequency_others %>% ungroup() %>% arrange(desc(counts)) %>% slice_max(counts)

tmp_df %>% 
  left_join(., y = frequency_others, by = "combination") %>%
  ggplot(mapping = aes(x = n_P, y = n_E, color = counts)) +
  geom_point() +
  coord_fixed(xlim = c(0,60), ylim = c(0,60)) + theme_light() + ggtitle("nonHK-CPs", subtitle = most_frequent$combination) + scale_color_gradient2( low= blue, mid = "gray", high=red)
```

```{r}
tmp_df <- plot_df %>% filter(is_HKG) %>% mutate(combination = paste(n_P, n_E, sep = "_"))
frequency_others <- tmp_df %>% group_by(combination) %>% count() %>% arrange(desc(n)) %>% rename(counts = n )
most_frequent <- frequency_others %>% ungroup() %>% arrange(desc(counts)) %>% slice_max(counts)

tmp_df %>% 
  left_join(., y = frequency_others, by = "combination") %>%
  ggplot(mapping = aes(x = n_P, y = n_E, color = counts)) +
  geom_point() +
  coord_fixed(xlim = c(0,60), ylim = c(0,60)) + theme_light() + ggtitle("HKG", subtitle = most_frequent$combination) + scale_color_gradient2( low= blue, mid = "gray", high=red)
```

```{r}
tmp_df <- plot_df %>% filter(is_HK_CP) %>%
  filter(!is_HKG) %>% mutate(combination = paste(n_P, n_E, sep = "_"))
frequency_others <- tmp_df %>% group_by(combination) %>% count() %>% arrange(desc(n)) %>% rename(counts = n )
most_frequent <- frequency_others %>% ungroup() %>% arrange(desc(counts)) %>% slice_max(counts)

tmp_df %>% 
  left_join(., y = frequency_others, by = "combination") %>%
  ggplot(mapping = aes(x = n_P, y = n_E, color = counts)) +
  geom_point() +
  coord_fixed(xlim = c(0,60), ylim = c(0,60)) + theme_light() + ggtitle("nonHKG", subtitle = most_frequent$combination) + scale_color_gradient2( low= blue, mid = "gray", high=red)
```


```{r}
#housekeeping core promoters
plot_df %>% filter(is_HK_CP) %>%
  ggplot(mapping = aes(y = p_P, x = p_E)) +
  geom_point(alpha = 0.1) + coord_fixed(xlim = c(0,1), ylim = c(0,1))
#non-housekeeping core promoters
plot_df %>% filter(!is_HK_CP) %>%
  ggplot(mapping = aes(y = p_P, x = p_E)) +
  geom_point(alpha = 0.1) + coord_fixed(xlim = c(0,1), ylim = c(0,1))
#housekeeping genes
plot_df %>% filter(is_HKG) %>%
  ggplot(mapping = aes(y = p_P, x = p_E)) +
  geom_point(alpha = 0.1) + coord_fixed(xlim = c(0,1), ylim = c(0,1))
#housekeeping core promoters, except housekeeping genes
plot_df %>% filter(is_HK_CP) %>%
  filter(!is_HKG) %>%
  ggplot(mapping = aes(y = p_P, x = p_E)) +
  geom_point(alpha = 0.1) + coord_fixed(xlim = c(0,1), ylim = c(0,1))
```

### Number of HK-CPs

```{r}
my_comparison <- list(c("Other", "HKG"), 
                      c("Other", "nonHKG"),
                      c("HKG", "nonHKG"))

p1 <- plot_df %>% 
  # filter(type_target != "Other") %>% 
  MyComparisonPlot(wt = "per_HK_CREs", 
                   compare_wt = "type_target", 
                   palette =  c("gray", red, blue), 
                   adjust = 1,
                   comparisons = my_comparison,
                   label = "p.signif",
                   vjust = 0.3,
                   box_width = 0.15) + 
  ylab(expression("Proportion of HK-CREs")) + 
  NoLegend() +
  theme(axis.title.x = element_blank(), plot.title = element_blank()) +
  theme(text = element_text(size = text_size)) +
  geom_jitter(size = 0.1, color = "black", alpha = 0.05)

p2 <- plot_df %>% 
  # filter(type_target != "Other") %>% 
  MyComparisonPlot(wt = "n_HK_CREs", 
                   compare_wt = "type_target", 
                   palette =  c("gray", red, blue), 
                   adjust = 1,
                   comparisons = my_comparison,
                   label = "p.signif",
                   vjust = 0.3,
                   box_width = 0.15) + 
  ylab(expression("Number of HK-CREs")) + 
  NoLegend() +
  theme(axis.title.x = element_blank(), plot.title = element_blank()) +
  theme(text = element_text(size = text_size)) +
  geom_jitter(size = 0.1, color = "black", alpha = 0.05)

p1 + p2

```

```{r}
#statistics other
plot_df %>% filter(type_target == "Other") %>% pull(n_HK_CREs) %>% summary()
#statistics HKG
plot_df %>% filter(type_target == "HKG") %>% pull(n_HK_CREs) %>% summary()
#statistics other
plot_df %>% filter(type_target == "nonHKG") %>% pull(n_HK_CREs) %>% summary()
```

```{r}
#statistics other
plot_df %>% filter(type_target == "Other") %>% pull(per_HK_CREs) %>% summary()
#statistics HKG
plot_df %>% filter(type_target == "HKG") %>% pull(per_HK_CREs) %>% summary()
#statistics other
plot_df %>% filter(type_target == "nonHKG") %>% pull(per_HK_CREs) %>% summary()
```

### Number of CTS-CPs

```{r}
my_comparison <- list(c("Other", "HKG"), 
                      c("Other", "nonHKG"),
                      c("HKG", "nonHKG"))

p1 <- plot_df %>% 
  MyComparisonPlot(wt = "per_CTS_CREs", 
                   compare_wt = "is_HK_CP", 
                   palette =  c("gray", red, blue), 
                   adjust = 2.5,
                   comparisons = my_comparison,
                   label = "p.signif",
                   vjust = 0.3,
                   box_width = 0.15) + 
  ylab(expression("Proportion of CTS-CREs")) + 
  NoLegend() +
  theme(axis.title.x = element_blank(), plot.title = element_blank()) +
  theme(text = element_text(size = text_size)) +
  geom_jitter(size = 0.1, color = "black", alpha = 0.05)

p2 <- plot_df %>% 
  MyComparisonPlot(wt = "n_CTS_CREs",  
                   compare_wt = "is_HK_CP", 
                   palette =  c("gray", red, blue), 
                   adjust = 2.5,
                   comparisons = my_comparison,
                   label = "p.signif",
                   vjust = 0.3,
                   box_width = 0.15) + 
  ylab(expression("Number of CTS-CREs")) + 
  NoLegend() +
  theme(axis.title.x = element_blank(), plot.title = element_blank()) +
  theme(text = element_text(size = text_size)) +
  geom_jitter(size = 0.1, color = "black", alpha = 0.05)

p1 + p2
```


```{r}
#statistics other
plot_df %>% filter(type_target == "Other") %>% pull(n_CTS_CREs) %>% summary()
#statistics HKG
plot_df %>% filter(type_target == "HKG") %>% pull(n_CTS_CREs) %>% summary()
#statistics other
plot_df %>% filter(type_target == "nonHKG") %>% pull(n_CTS_CREs) %>% summary()
```

```{r}
#statistics other
plot_df %>% filter(type_target == "Other") %>% pull(per_CTS_CREs) %>% summary()
#statistics HKG
plot_df %>% filter(type_target == "HKG") %>% pull(per_CTS_CREs) %>% summary()
#statistics other
plot_df %>% filter(type_target == "nonHKG") %>% pull(per_CTS_CREs) %>% summary()
```










