---
title: "Analysis average HiC and housekeeping core promoters"
author: "Martin Loza"
date: "2023/12/19
---

On this workflow I want to analyse the overlapping between HK-CPs and the average HiC. I want to confirm if these elements have more interactions as compared with CTS-CREs

```{r setup}
library(dplyr)
library(ggplot2)
library(patchwork)
library(GenomicRanges)
library(ggpubr)

## Global variables
in_dir  <- "~/Documents/Projects/HK_Interactions/Analysis/2023_12/2023_12_19/Results/"
out_dir <- "~/Documents/Projects/HK_Interactions/Analysis/2023_12/2023_12_19/Results/"

date = 231219
red = "#CC6677"
blue = "#6699CC"

#Functions
source(file = "~/Projects/HK_Interactions/Analysis/Functions.R")
source(file = "~/Projects/HK_Interactions/Analysis/Functions_Visualizations.R")

#Local functions 
MyComparisonPlot <- function(.data = NULL, wt = NULL, box_width = 1,                             compare_wt = NULL, palette = NULL, adjust = 1, ...){
  lim_y = c(min(.data[[wt]]), max(.data[[wt]]) + 1.5)
  
  p <- .data %>% 
    MyViolinPlot(x = compare_wt, y = wt, fill_violin = compare_wt, adjust = adjust)
  
  if(!is.null(palette)){
    p <- p + scale_fill_manual(values = palette)
  }
  
  p <- p + geom_boxplot(fill = "white", width = box_width,
                        linewidth = 0.3, outlier.size = 0.01, outlier.color = "gray") +
    ggtitle(wt) + 
    ggpubr::stat_compare_means(method = "wilcox.test", paired = FALSE, ...) 
  
  return(p)
}
```

## Load average data

```{r}
resolution = '5000'
average_data = read.csv(file = paste0("~/Projects/HK_Interactions/Data/HiC/4DN_portal/average/01_bedpe/average_matrix_", resolution, ".bed"), 
                        header = FALSE, sep = "\t", )
average_data <- as.data.frame(average_data)
colnames(average_data) <- c("chr1","start1", "end1", "chr2", "start2", "end2", "mean_contact")
```

Load annotated interactions 

```{r}
# interactions <- readRDS("/Volumes/MARTIN_LOZA/Projects/HK_Interactions/Data/Commonly_used/interactions_all_230912.rds")
# interactions %>% filter(!duplicated(region_id)) %>% pull(n_celltypes) %>% table()
```

```{r}
regions <- readRDS("~/Projects/HK_Interactions/Data/HK_CREs/ns/annotated_regions_all_230525.rds")

regions <- regions %>% 
  mutate(test_n_celltypes = factor(test_n_celltypes))
levels(regions$test_n_celltypes) <- c("NS", 1:50)
```


## Initial tests and setup

Only intra-interactions?

```{r}
sum(average_data$chr1 != average_data$chr2) == 0
```

Setup average data

```{r}
average_data$chr1 <- factor(average_data$chr1, levels = paste0("chr", c(1:22, "X")))
average_data$chr2 <- factor(average_data$chr2, levels = paste0("chr", c(1:22, "X")))
average_data$start1 <- as.integer(average_data$start1)
average_data$end1 <- as.integer(average_data$end1)
```

```{r}
#remove diagonal
average_data <- average_data %>% mutate(id1 = paste(chr1,":",start1,"-",end1),
                                      id2 = paste(chr2,":",start2,"-",end2))
average_data <- average_data %>% filter(id1 != id2)
```


Get only CREs, e.g. we don't care about their interaction

```{r}
# cres <- interactions %>% GetUniqueRegions()
cres <- regions 
table(cres$test_n_celltypes)
```
Setup Cres

```{r}
cres$start <- as.integer(cres$start)
cres$end <- as.integer(cres$end)
```

## Analysis

I would like to get the average pixels related to each CTS-CREs and HK-CREs

```{r}
#let's focus on CTS and HK
sel_cres <- cres %>% filter(n_celltypes == 0 | n_celltypes == 1 | n_celltypes == 50)
table(sel_cres$n_celltypes)

sel_cres <- sel_cres %>%
  mutate(n_celltypes = factor(n_celltypes, levels = c(0,1,50)))
levels(sel_cres$n_celltypes) <- c("NS", "CTS-CREs", "HK-CREs")
table(sel_cres$n_celltypes)
```

```{r}
# I need to get the bins that overlaps with sel_cres 
# I need to do this in a loop, to annotate each cres

#get only pixels related to sel_cres

#select columns of interest
sel_pixels <- average_data %>% select(chr1, start1, end1, mean_contact) 
#rename columns
colnames(sel_pixels) <- c("chr", "start", "end", "mean")

#get index of overlapping cres and pixes
ovl <- WhichRegionsOverlap(query_enhancer = sel_cres, subject_enhancer = sel_pixels)
sub_idx <- subjectHits(ovl)
que_idx <- queryHits(ovl)
# sel_pixels <- sel_pixels[unique(idx),]

sel_cres <- sel_cres %>% mutate(average_contacts = NA)
sel_cres <- sel_cres %>% mutate(n_contacts = 0)
```

```{r}
contact_thr = 1.0
#for each cre in sel_cres
for(i in seq_len(nrow(sel_cres))){
#i = 1
  #get ovl pixels with current cres
  c_idx <- which(que_idx == i)
  #get overlappin pixels
  ovl_idx <- sub_idx[c_idx]
  #get mean contacts
  c_mean_contacts <- sel_pixels[ovl_idx,"mean"]
  #remove contacts lower than threshold
  c_mean_contacts <- c_mean_contacts[c_mean_contacts>= contact_thr]
  #if not empty, calculate the average
  if(isEmpty(c_mean_contacts)){
    c_mean_contacts <- NA
    n_contacts <- 0
  }else{
    n_contacts <- length(c_mean_contacts)
    c_mean_contacts <- mean(c_mean_contacts)
  }
  #%>% mean()
  #assign the mean contacts
  sel_cres[i,"average_contacts"] <- c_mean_contacts
  #assign number of contacts
  sel_cres[i,"n_contacts"] <- n_contacts
}
```

```{r}
# save 
# saveRDS(object = sel_cres, file = paste0(out_dir,"cres_mean_values_", resolution, ".rds"))
# sel_cres <- readRDS(file = paste0(out_dir,"cres_mean_values_100000.rds"))
```



```{r}
my_comparison <- list(c("NS", "CTS-CREs"), c("NS", "HK-CREs"), c("CTS-CREs", "HK-CREs"))

p <- sel_cres %>% 
  MyComparisonPlot(wt = "average_contacts",
                   compare_wt = "n_celltypes", 
                   palette =  c("gray", red, blue), adjust = 1.7,
                   comparisons = my_comparison,
                   label = "p.signif", vjust = 0.5, box_width = 0.15) + 
  ylab(expression("Average HiC")) +
  theme(axis.title.x = element_blank(), plot.title = element_blank())

p

```
```{r}
my_comparison <- list(c("NS", "CTS-CREs"), c("NS", "HK-CREs"), c("CTS-CREs", "HK-CREs"))

p <- sel_cres %>% 
  mutate(log_n_contacts = log10(n_contacts+1)) %>%
  MyComparisonPlot(wt = "log_n_contacts",
                   compare_wt = "n_celltypes", 
                   palette =  c("gray", red, blue), adjust = 1.7,
                   comparisons = my_comparison,
                   label = "p.signif", vjust = 0.5, box_width = 0.15) + 
  ylab(expression("Number of contacts (log" [10] ~ ")")) +
  theme(axis.title.x = element_blank(), plot.title = element_blank()) +
  ggtitle(paste0("Resolution: ",resolution ))
  # geom_jitter(size = 0.001, color = "gray", alpha = 0.05)

p

```
```{r}
#boxplot by chromosome
sel_cres %>% 
  MyBoxPlot(x = "n_celltypes", y = "average_contacts", outlier.size = 0.5) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  ggtitle(paste0("Resolution:", resolution, ". Number of interactions:", nrow(average_data))) + 
  ylab("Mean(Scaled(ICE-balanced(counts)))")
```



























