---
title: "HK-PPI-hubs"
author: "Martin Loza"
date: "2023/10/17
---

I would like to analyze housekeeping PPI hubs

NOTES:


```{r setup}
library(dplyr)
library(ggplot2)
library(patchwork)
library(stringr)
library(GenomicRanges)
library(igraph)
library(ggraph)
library(readr)

## Global variables
in_dir  <- "~/Documents/Projects/HK_Interactions/Analysis/2023_09/2023_09_18/Results/"
out_dir <- "~/Documents/Projects/HK_Interactions/Analysis/2023_09/2023_09_18/Results/"

date = 231017

#Functions
source(file = "~/Documents/Projects/HK_Interactions/Analysis/Functions.R")
source(file = "~/Documents/Projects/HK_Interactions/Analysis/Functions_Visualizations.R")

#Local functions 

GetGenerationEdges <- function(edges = NULL, init_nodes = NULL, n_generations = 2){
  
  #loop. In each generation get the edges associated to init nodes
  for(i in 2:n_generations){
    tmp_df <- edges %>% filter(gene_from %in% all_of(init_nodes) | gene_to %in% all_of(init_nodes))
    init_nodes <- union(tmp_df$gene_from, tmp_df$gene_to)
  }

  return(tmp_df)
}
```

## Load data

Load annotated interactions 

```{r}
interactions <- readRDS(paste0(in_dir,"/interactions_all_230912.Rds"))
interactions %>% filter(!duplicated(region_id)) %>% pull(n_celltypes) %>% table()
```

## Set up

I would like to select interactions only for CP 

```{r}
interactions <- interactions %>% filter(ann_v1 == "CP")
interactions$ann_v1 <- droplevels(interactions$ann_v1)
interactions %>% filter(!duplicated(region_id)) %>% pull(n_celltypes) %>% table()
```

Check for duplicated interactions 

```{r}
any(duplicated(interactions$id))
```

Set the absolute distance of interactions 

```{r}
interactions <- interactions %>%
  mutate(abs_distance = abs(distance))
```

## Analysis

Let's select only those interactions found in more than 45 cell types

```{r}
hk_ppi <- interactions %>% filter(n_ct_interactions >= 45)
```

### Directed graph

For now, let's assume that the regions corresponds to an annotated gene
Later, let's check for divergent genes too. For now, let's use the nearest genes

We need to get a data frame with the edges and the weight information

```{r}
#set new genes with direction
hk_ppi <- hk_ppi %>% 
  mutate(gene_from = nearest_gene, gene_to = gene) %>%
  mutate(weight = n_ct_interactions/50) #normalized weights
#get the edges data frames
edges_df <- hk_ppi %>% dplyr::select(gene_from, gene_to, weight)
```

Get the directed graph

```{r}
#get graph from edges
graph_dir<- graph.data.frame(d = edges_df, directed = TRUE)
```

In the case of weighted graphs, the degree it's called strength. 
Let's get the strength of the nodes

```{r}
strength <- strength(graph = graph_dir, mode = "all", loops = TRUE)
strength <- data.frame(gene = names(strength), strength = strength)
```

Let's try to select the top nodes

```{r}
n_top = 1000
top_nodes <- strength %>% slice_max(order_by = strength, n = n_top) %>% arrange(desc(strength))
hub_genes <- top_nodes$gene
```

### setup hubs 

I would like to analyze the core promoters of hub genes in the micro-c data

let's get the ppi of hub genes

```{r}
# let's focus on the ePromoters, promoters with enhancer capabilities
hub_ppi <- hk_ppi %>% filter(gene_from %in% hub_genes)
```

let's check how many regions we got 

```{r}
#number of unique regions
hub_ppi %>% filter(!duplicated(region_id)) %>% nrow()
```

we have more regions than unique hubs. This is because a gene hub could have more than one associated TSS. Therefore, for each gene, let's focus on the transcript with higher number of interactions.

```{r}
#get the top TSS of each hub
hub_region <- hub_ppi %>% 
  group_by(gene_from) %>%
  count(region_id) %>%
  slice_max(n, n = 1) %>% 
  ungroup() %>% select(region_id)

cat("Number of regions: ", nrow(hub_region))
```

### Top Hubs

```{r}
#fix length
l = 3000

#select center and chromosome of hub regions.
#NOTE: there can be more than one region than gene_from as they can have the same number of target genes, so more than one regions is selected after slice_max
sel_regions <- hk_ppi %>% 
  filter(!duplicated(region_id)) %>% 
  filter(region_id %in% hub_region$region_id) %>% 
  select(chr, center) %>% 
  mutate(start = center - l, end = center + l)
```

```{r}
#setup temporal bedpe
tmp_bed <- sel_regions %>%  
  select(chr, start, end)
table(tmp_bed$chr)

#save bedpe
write.table(x = tmp_bed,
            paste0(file = "~/Documents/Projects/HK_Interactions/Analysis/2023_10/2023_10_17/tmp_results/hub_ppi_top_",n_top,"_3k.bed"),
            quote = FALSE, sep = "\t", row.names = FALSE, col.names = FALSE)
```

### Other HK-PPI

### CTS-CP



### Load results from pscan analysis

```{r}
hk_ppi_tf <- readr::read_table(file = "~/Documents/Projects/HK_Interactions/Analysis/2023_10/2023_10_17/tmp_results/hub_ppi_top_1000_TF.txt", col_names = TRUE)
```

set up 

```{r}
#selected columns
sel_cols <- c("NAME", "ID", "L_PVALUE", "O/U", "G_PVALUE", "O/U_1")
hk_ppi_tf <- hk_ppi_tf %>% select(all_of(sel_cols))
#rename the columns
colnames(hk_ppi_tf) <- c("NAME", "ID", "L_PVALUE", "L_O_U", "G_PVALUE", "G_O_U")
```

convert character pvalues to integer

```{r}
hk_ppi_tf$L_PVALUE <- as.numeric(hk_ppi_tf$L_PVALUE)
hk_ppi_tf$G_PVALUE <- as.numeric(hk_ppi_tf$G_PVALUE)
```

### TOP GLOBAL OVER and UNDER 

```{r}
th = 0.005
top_over <- hk_ppi_tf %>% filter(G_PVALUE <= th & G_O_U == "OVER") %>% arrange((G_PVALUE))
top_under <- hk_ppi_tf %>% filter(G_PVALUE <= th & G_O_U == "UNDER") %>% arrange((G_PVALUE))
```

```{r}
top_over %>% arrange(NAME) %>% pull(NAME)
```

```{r}
top_under %>% arrange(G_PVALUE) %>% slice_head(n = 100) %>%  arrange(NAME) %>% pull(NAME)
```

```{r}
top_under %>% arrange(G_PVALUE) %>% slice_head(n = 100) %>%  arrange(NAME) %>% pull(NAME)
```





















